# Project: Patient Preferred Language And Deep Sedation During Invasive Mechanical Ventilation: A Multicenter Retrospective Cohort Study
# Author: Alex Ortiz

# PART ZERO: Setup Step - Setup libraries
```{r}
# Install Packages

packages <- c("zoo", "data.table", "dplyr", "lubridate", "tidyr", "fst", "arrow", "table1", "lme4", "metafor", "yaml", "pscl", "MASS", "htmltools", "car", "WeightIt", "MatchIt", "purrr", "survival", "mediation", "cmprsk", "gfoRmula", "nnet", "medflex", "bruceR", "tidyverse", "survey", "ipw", "geepack", "readr", "collapse", "stringr", "rprojroot", "jsonlite")
install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}
 
sapply(packages, install_if_missing)

# Load necessary libraries
library(data.table)
library(dplyr)
library(zoo)  # for na.locf function
library(lubridate)
library(tidyr)
library(fst)
library(arrow)
library(table1)
library(metafor)
library(yaml)
library(htmltools)
library(MASS)
library(pscl)
library(car)
library(stringr)
library(rprojroot)
library(jsonlite)

# Read the YAML or JSON config from the config folder
project_root <- file.path( find_root(rprojroot::has_dir("CLIF_Deep_Sedation_Language")),
  "CLIF_Deep_Sedation_Language")
knitr::opts_knit$set(root.dir = file.path(project_root, "config"))
config <- yaml::read_yaml(file.path(project_root, "config", "CLIF_Deep_Sedation_Language_config.yaml"))

#config <- jsonlite::fromJSON("config/CLIF_Deep_Sedation_Language_config.json")

# Define directories based on YAML configuration
data_dir <- config$data_dir
output_dir <- config$output_dir
intermediate_dir <- file.path(output_dir, "intermediate_data/")  # Define output_dir based on data_dir
tables_dir <- file.path(output_dir, "final_tables/")  
final_data <- file.path(output_dir, "final_data/")
second_dir <- file.path(output_dir, "secondary_project/")
file_type <- config$file_type
institution <- config$institution
time_zone <- config$time_zone

dirs_to_create <- c(intermediate_dir, tables_dir, final_data, second_dir)
for (dir in dirs_to_create) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)
  }
}
```

# PART ONE: Upload CLIF Respiratory, Hospitalization, ADT, Patient Tables
```{r}
read_data <- function(file_path, select = NULL) {
  if (grepl("\\.csv$|\\.csv\\.gz$", file_path)) {
    # fread handles select = NULL by loading all columns
    return(fread(file_path, select = select))
  } else if (grepl("\\.parquet$", file_path)) {
    # Use col_select only if select is not NULL
    if (is.null(select)) {
      return(arrow::read_parquet(file_path))
    } else {
      return(arrow::read_parquet(file_path, col_select = all_of(select)))
    }
  } else if (grepl("\\.fst$", file_path)) {
    # Use columns only if select is not NULL
    if (is.null(select)) {
      return(fst::read_fst(file_path))
    } else {
      return(fst::read_fst(file_path, columns = all_of(select)))
    }
  } else {
    stop("Unsupported file format. Only CSV, Parquet, and FST are supported.")
  }
}


#Load Raw CLIF Respiratory Support, Hospitalization, Patient, ADT Tables
respiratory_support <- as.data.table(read_data(paste0(data_dir, "clif_respiratory_support", file_type), select = c("hospitalization_id", "recorded_dttm", "device_category", "mode_category", "tracheostomy", "fio2_set", "peep_set", "lpm_set", "tidal_volume_set")))
respiratory_support <- unique(respiratory_support, by = c("hospitalization_id", "recorded_dttm"))

hospitalization <- as.data.table(read_data(paste0(data_dir, "clif_hospitalization", file_type)))
hospitalization <- unique(hospitalization, by = "hospitalization_id")
#Filter Out Incomplete Hospitalizations
hospitalization <- hospitalization[!is.na(discharge_dttm)]

patient_demographics <- as.data.table(read_data(paste0(data_dir,"clif_patient", file_type), select = c("patient_id", "sex_category", "race_category", "ethnicity_category", "death_dttm", "language_category")))
patient_demographics <- unique(patient_demographics, by = "patient_id")

adt <- as.data.table(read_data(paste0(data_dir, "clif_adt", file_type), select = c("hospitalization_id", "hospital_id", "in_dttm", "out_dttm", "location_category", "location_type")))
setorder(adt, hospitalization_id, in_dttm)

```

# PART TWO: Identify IMV Runs of at least 24 hours
```{r}
# Identify All Hospitalizations with IMV
resp_supp_imv_id <- respiratory_support[device_category == "IMV"]

imv_hospitalizations <- hospitalization[hospitalization_id %in% resp_supp_imv_id$hospitalization_id]

# Total Respiratory Support Table for IMV Hospitalizations
respiratory_support_imv <- respiratory_support[hospitalization_id %in% resp_supp_imv_id$hospitalization_id]
setorder(respiratory_support_imv, hospitalization_id, recorded_dttm)

# Convert blank values in device_category to "NA"
respiratory_support_imv[device_category == "", device_category := NA]

# Identify IMV Runs According to Rule - Two consecutive IMV values in device_category to start IMV time and two consecutive non-IMV values to end IMV time. If end IMV time is not IDed by algorithm then use recorded_dttm from last row (patient likely discharged - either transferred or deceased)
imv_runs <- respiratory_support_imv[, {
  n <- .N
  run_list <- list()
  i <- 1
  
  while (i <= n) {
    current <- device_category[i]
    
    # Look ahead for second IMV within next few rows (skipping NAs)
    found_second_imv <- FALSE
    for (k in (i + 1):n) {
      if (!is.na(device_category[k]) && device_category[k] == "IMV") {
        found_second_imv <- TRUE
        break
      } else if (!is.na(device_category[k]) && device_category[k] != "IMV") {
        break  # Stop looking if we hit a non-NA, non-IMV
      }
    }
    
    if (!is.na(current) && current == "IMV" && found_second_imv) {
      begin_idx <- i
      begin_imv <- recorded_dttm[begin_idx]
      
      # Search for end: 2 consecutive non-IMV/non-NA rows
      end_idx <- NA
      trach_flag <- FALSE
      for (j in (begin_idx + 1):(n - 1)) {
        dj1 <- device_category[j]
        dj2 <- device_category[j + 1]
        
        cond1 <- !is.na(dj1) && dj1 != "IMV"
        cond2 <- !is.na(dj2) && dj2 != "IMV"
        
        if (cond1 && cond2) {
          end_idx <- j
          if (dj1 == "Trach Collar" || dj2 == "Trach Collar") {
            trach_flag <- TRUE
          }
          break
        }
      }
      
      # Use last timestamp if no end found
      if (is.na(end_idx)) {
        end_imv <- recorded_dttm[n]
      } else {
        end_imv <- recorded_dttm[end_idx]
      }
      
      run_list[[length(run_list) + 1]] <- data.table(
        hospitalization_id = hospitalization_id[1],
        begin_imv = begin_imv,
        end_imv = end_imv
      )
      
      if (trach_flag) break
      
      i <- if (!is.na(end_idx)) end_idx + 1 else n + 1
    } else {
      i <- i + 1
    }
  }
  
  rbindlist(run_list)
}, by = hospitalization_id]


# Add duration of IMV and run_id to filter for first period of IMV
imv_runs[, `:=`(
  imv_duration_hours = as.numeric(difftime(end_imv, begin_imv, units = "hours")),
  run_id = seq_len(.N)
), by = hospitalization_id]

#Reintubations
reintubations <- imv_runs[run_id >= 2]

# Only Retain the first IMV Run for each hospitalization_id 
first_imv_runs <- imv_runs[, .SD[1], by = hospitalization_id]

imv_hospitalizations_2 <- imv_hospitalizations[hospitalization_id %in% first_imv_runs$hospitalization_id]
first_imv_runs_final <- merge(first_imv_runs, imv_hospitalizations_2, by = "hospitalization_id", all = TRUE)

#Correct IMV Times for Those with end_imv times after discharge_dttm and then re-calculate the duration and remove anybody less than 24 hours
first_imv_runs_final[
  end_imv > discharge_dttm, end_imv := discharge_dttm
]

first_imv_runs_final[, `:=`(
  imv_duration_hours = as.numeric(difftime(end_imv, begin_imv, units = "hours"))
), by = hospitalization_id]

# Filter rows where imv_duration_hours is greater than or equal to 24 hours
part_two_imv_runs <- first_imv_runs_final[imv_duration_hours >= 24] 
part_two_imv_runs <- part_two_imv_runs[, .(hospitalization_id, begin_imv, end_imv, imv_duration_hours)]
```

# PART THREE: Identify Trach Patients and Remove Them from The Cohort
```{r}
# Identify hospitalization_ids that have tracheostomy at some point during their hospitalization and earliest timestamp
trach_resp <- respiratory_support[tracheostomy == "1"]
trach_dttm_table <- trach_resp[, .(trach_dttm = min(recorded_dttm)), by = hospitalization_id]
hosp_id_trach <- unique(trach_resp$hospitalization_id)

# Identify hospital_ids that have "trach collar" device at any point in hospitalization and corresponding earliest timestamp
trach_collar_resp <- respiratory_support[device_category == "Trach Collar"]
trach_collar_dttm_table <- trach_collar_resp[, .(trach_collar_dttm = min(recorded_dttm)), by = hospitalization_id]
hosp_id_trach_collar <- unique(trach_collar_resp$hospitalization_id)

# Merge the Two Trach Hospital DTTM Tables
trach_times <- merge(trach_dttm_table, trach_collar_dttm_table, by = "hospitalization_id", , all = TRUE)
trach_times[, first_trach_dttm := pmin(trach_dttm, trach_collar_dttm, na.rm = TRUE)]

# Hospitalization Table for Trach Hospitalization_IDs
hosp_id_trach_total <- unique(trach_times$hospitalization_id)
hospitalization_trach <- hospitalization[hospitalization_id %in% hosp_id_trach_total]

# Pull all hospitalizations for each patient_id associated with a trach hospitalization
all_hospitalizations_trach_patients <- hospitalization[patient_id %in% hospitalization_trach$patient_id]

# Merge Hospitalization and Trach Times and perform LOCF to fill first_trach_dttm
all_hosp_w_trach_times <- merge(all_hospitalizations_trach_patients, trach_times, by = "hospitalization_id", , all = TRUE)
setorder(all_hosp_w_trach_times, patient_id, admission_dttm)
all_hosp_w_trach_times[, first_trach_dttm := na.locf(first_trach_dttm, na.rm = FALSE), by = patient_id]

# Since we have filled out first_trach_dttm for all hospitalizations for patients with any history of trach we can now filter the table only for those with >24-hour IMV runs
imv_pts_w_trach <- all_hosp_w_trach_times[hospitalization_id %in% part_two_imv_runs$hospitalization_id]
setorder(imv_pts_w_trach, patient_id, admission_dttm)

# Merge Trach IMV Patients with IMV Run Details
full_trach_table <- merge(imv_pts_w_trach, part_two_imv_runs, by = "hospitalization_id", , all = TRUE)
full_trach_table_filt <- full_trach_table[hospitalization_id %in% imv_pts_w_trach$hospitalization_id]
full_trach_table_filt <- full_trach_table_filt[, .(hospitalization_id, patient_id, admission_dttm, discharge_dttm, discharge_category, begin_imv, end_imv, first_trach_dttm)]

# Create a Flag for Being Trached During Admission
full_trach_table_filt[, trached_during_admission := 
  !is.na(first_trach_dttm) & 
  first_trach_dttm >= (admission_dttm - lubridate::days(1)) & 
  first_trach_dttm <= (discharge_dttm + lubridate::days(1))
]

# Create a Flag for Being Trached in the 6 Months Prior to Admission
full_trach_table_filt[, trached_within_6months_before_admission := 
  !trached_during_admission & 
  !is.na(first_trach_dttm) & 
  (difftime(admission_dttm, first_trach_dttm, units = "days") <= 6 * 30.44) & 
  (difftime(admission_dttm, first_trach_dttm, units = "days") > 0)
]

# Create a Flag for Being Trached within 72 hours of IMV
full_trach_table_filt[, trached_within_72_hours_of_imv := 
  trached_during_admission & 
  !is.na(begin_imv) & 
  first_trach_dttm <= (begin_imv + lubridate::hours(72))
]

# Create a Flag for Missed Trach Patients - Those Still Intubated within 2 Hours of Discharge who are Discharged to something other than "Expired", "Other", or "Acute Care Hospital"
full_trach_table_filt[, ventilated_at_discharge := ifelse(
  difftime(discharge_dttm, end_imv, units = "hours") <= 2 | end_imv > discharge_dttm,
  TRUE, FALSE
)]
full_trach_table_filt[, missed_trach := 
  ifelse(ventilated_at_discharge == "TRUE" & 
         !(discharge_category %in% c("Expired", "Other", "Acute Care Hospital", "Hospice")), 
         TRUE, FALSE)]

#Trach Hospitalization_IDs to remove
trach_hospitalizations_remove <- full_trach_table_filt[
  trached_within_72_hours_of_imv == TRUE | trached_within_6months_before_admission == TRUE | missed_trach == TRUE, 
  .(hospitalization_id)
]

#Remove the trach only hospitalizations from the imv runs
imv_runs_trach_filt <- part_two_imv_runs[!hospitalization_id %in% trach_hospitalizations_remove$hospitalization_id]

trach_times_filt <- trach_times[hospitalization_id %in% imv_runs_trach_filt$hospitalization_id]

imv_runs_trach_filtered <- merge(imv_runs_trach_filt, trach_times_filt, by = "hospitalization_id", , all = TRUE)

imv_runs_trach_filtered <- imv_runs_trach_filtered[, .(hospitalization_id, begin_imv, end_imv, first_trach_dttm, imv_duration_hours)]

#Reintubation Information on Cleaned Hospitalization_IDs
reintubations_filtered <- reintubations[hospitalization_id %in% imv_runs_trach_filtered$hospitalization_id]
reintubations_filtered[, reintubation := ifelse(run_id >= 2, 1, 0)]

reintubations_filtered <- reintubations_filtered[, .(hospitalization_id, reintubation)]

#Put together part_three imv runs combining trach information, reintubation information
part_three_imv_runs <- merge(imv_runs_trach_filtered, reintubations_filtered, by = "hospitalization_id", , all = TRUE)

part_three_imv_runs[is.na(reintubation), reintubation := 0]

#Need to Calculate imv_duration_hours_trach variable for non re-intubated patients
part_three_imv_runs[, imv_duration_hours_trach := 
  ifelse(reintubation == 0, 
         as.numeric(difftime(first_trach_dttm, begin_imv, units = "hours")), 
         NA_real_), 
  by = hospitalization_id]

#Trached Flag
part_three_imv_runs[, trached := fifelse(!is.na(first_trach_dttm), 1, 0)]

part_three_imv_runs <- unique(part_three_imv_runs, by = c("hospitalization_id"))
```

# PART FOUR: Remove OSH Transfers Intubated at Admission
```{r}
first_device <- merge(respiratory_support, hospitalization, by = "hospitalization_id", all.x = TRUE)
first_device_filt <- first_device[hospitalization_id %in% part_three_imv_runs$hospitalization_id]
#Remove NA device_category rows
first_device_filt <- first_device_filt[!is.na(device_category)]
setorder(first_device_filt, hospitalization_id, recorded_dttm)

earliest_device <- first_device_filt[, .SD[1], by = hospitalization_id, .SDcols = c("recorded_dttm", "device_category")]
setnames(earliest_device, "device_category", "first_device_category")
setnames(earliest_device, "recorded_dttm", "first_device_dttm")

first_device_final <- merge(hospitalization, earliest_device, by = "hospitalization_id", all.x = TRUE)
first_device_final_filt <- first_device_final[hospitalization_id %in% part_three_imv_runs$hospitalization_id]

intubated_transfers <- first_device_final_filt[
  first_device_category == "IMV" & admission_type_category %in% c("osh", "facility", "OSH")
]

intubated_transfers[, time_diff := as.numeric(difftime(first_device_dttm, admission_dttm, units = "hours"))]

part_four_imv_runs <- part_three_imv_runs[!hospitalization_id %in% intubated_transfers$hospitalization_id]
```

# PART FIVE: Determine Intubation Times, Primary ICU, Hospital ID
```{r}
adt_filt_icu <- adt[hospitalization_id %in% part_four_imv_runs$hospitalization_id]
adt_icu <- merge(part_four_imv_runs, adt_filt_icu, by = "hospitalization_id", all.x = TRUE)

#Determine Location of Intubation
# Step 1: Try to find the row where begin_imv falls between in_dttm and out_dttm
intubation_location <- adt_icu[
  begin_imv >= in_dttm & begin_imv <= out_dttm,
  .SD[which.min(abs(difftime(begin_imv, in_dttm)))],  # use closest if multiple
  by = hospitalization_id
][
  , .(hospitalization_id, location_of_intubation = location_category)
]

# Step 2: Identify fallback rows where begin_imv is earlier than all in_dttm
fallback_location <- adt_icu[
  , .SD[which.min(in_dttm)],  # first row per hospitalization
  by = hospitalization_id
][
  begin_imv < in_dttm,
  .(hospitalization_id, location_of_intubation = location_category)
]

# Step 3: Combine results, prioritizing primary case, using coalesce to fill missing
all_location_of_intubation <- rbind(intubation_location, fallback_location)[
  order(hospitalization_id)
][
  , .SD[1], by = hospitalization_id  # take only one row per ID
]

#For Procedural Intubations Determine Length of Procedure
# Step 1: Get hospitalization_ids with 'procedural' location_of_intubation
procedural_ids <- all_location_of_intubation[location_of_intubation == "procedural", hospitalization_id]

# Step 2: Subset adt_icu to just those IDs
adt_icu_procedural <- adt_icu[hospitalization_id %in% procedural_ids]

# Step 3: Find the intubation row for each patient
intubation_rows <- adt_icu_procedural[
  begin_imv >= in_dttm & begin_imv <= out_dttm,
  .SD[which.min(abs(difftime(begin_imv, in_dttm)))],  # primary match
  by = hospitalization_id
]

# Step 4: Add fallback rows if none match (begin_imv < all in_dttm)
fallback_rows <- adt_icu_procedural[
  , .SD[which.min(in_dttm)],
  by = hospitalization_id
][
  begin_imv < in_dttm
]

# Step 5: Combine and de-duplicate
intubation_rows_combined <- rbind(intubation_rows, fallback_rows)[
  order(hospitalization_id)
][
  , .SD[1], by = hospitalization_id
]

# Step 6: Calculate procedure_time in hours
intubation_rows_combined[, procedure_time := as.numeric(difftime(out_dttm, in_dttm, units = "hours"))]

# Result: hospitalization_id and procedure_time for procedural intubations
procedure_duration_table <- intubation_rows_combined[, .(hospitalization_id, procedure_time)]

# Merge procedure duration into all_location_of_intubation
all_location_of_intubation <- merge(
  all_location_of_intubation,
  procedure_duration_table,
  by = "hospitalization_id",
  all.x = TRUE
)

#Determine Time from Intubation to ICU Admission for non-procedural and non-ICU intubations
# Step 1: Filter out ICU and procedural intubations
non_icu_procedural <- all_location_of_intubation[!location_of_intubation %in% c("icu", "procedural")]

# Step 2: Get first ICU admission time after intubation for these patients
icu_admission_times <- adt_icu[
  hospitalization_id %in% non_icu_procedural$hospitalization_id &
    location_category == "icu" &
    in_dttm >= begin_imv,
  .SD[which.min(in_dttm)],  # get first ICU entry after intubation
  by = hospitalization_id
][, .(hospitalization_id, begin_imv, icu_admission_time = in_dttm)]

# Step 3: Merge ICU admission times into the non_icu_procedural table
non_icu_procedural <- merge(non_icu_procedural, icu_admission_times, by = "hospitalization_id", all.x = TRUE)

# Step 4: Calculate time from intubation to ICU admission in hours
non_icu_procedural[, time_to_icu_admission := as.numeric(difftime(icu_admission_time, begin_imv, units = "hours"))]

#Merge Back Together and Filter
intubation_icu_timeline <- merge(
  all_location_of_intubation,
  non_icu_procedural[, .(hospitalization_id, icu_admission_time, time_to_icu_admission)],
  by = "hospitalization_id",
  all.x = TRUE
)

#Clean Out Procedures Longer than 4 hours
intubation_icu_timeline_clean_procedure <- intubation_icu_timeline[is.na(procedure_time) | procedure_time <= 4]

#Clean Out ICU Transfers Longer than 4 hours
intubation_icu_timeline_clean <- intubation_icu_timeline_clean_procedure[is.na(time_to_icu_admission) | time_to_icu_admission <= 4]

#Final Intubation Table
location_of_intubation_final <- intubation_icu_timeline_clean[, .(hospitalization_id, location_of_intubation, time_to_icu_admission)]

####

adt_icu_clean <- merge(adt_icu, location_of_intubation_final, by = "hospitalization_id")

# Find first ICU after beginning IMV to determine primary ICU
first_icu_post_imv <- adt_icu_clean[
  location_category == "icu" &
    (in_dttm > begin_imv | (begin_imv >= in_dttm & begin_imv <= out_dttm)),
  .SD[which.min(in_dttm)],
  by = hospitalization_id
]

# Find first location_category by in_dttm for each hospitalization_id in the filtered set
first_location <- adt_icu_clean[
  hospitalization_id %in% first_icu_post_imv$hospitalization_id & 
    !is.na(location_category),
  .SD[which.min(in_dttm)],
  by = hospitalization_id
][
  , .(hospitalization_id, first_location_category = location_category)
]

# Merge back into first_icu_post_imv_filt
first_icu_post_imv_filt <- merge(first_icu_post_imv, first_location, by = "hospitalization_id", all.x = TRUE)

# Set names
setnames(first_icu_post_imv_filt, "first_location_category", "first_location_in_hospital")
setnames(first_icu_post_imv_filt, "location_type", "icu_type")

set(first_icu_post_imv_filt, j = "hospital_icu_id", 
    value = paste(first_icu_post_imv_filt$hospital_id, 
                  first_icu_post_imv_filt$icu_type, sep = "_"))

first_icu_post_imv_final <- first_icu_post_imv_filt[, .(hospitalization_id, hospital_id, icu_type, hospital_icu_id, location_of_intubation, first_location_in_hospital, time_to_icu_admission)]

#Clean out any hospitalizations with missing data
first_icu_post_imv_final <- first_icu_post_imv_final[!is.na(icu_type)]
first_icu_post_imv_final <- first_icu_post_imv_final[!is.na("hospital_id")]

# Create part five IMV Runs
xx_part_four_filt <- part_four_imv_runs[hospitalization_id %in% first_icu_post_imv_final$hospitalization_id]

part_five_imv_runs <- merge(xx_part_four_filt, first_icu_post_imv_final, by = "hospitalization_id", all = TRUE)
part_five_imv_runs <- unique(part_five_imv_runs, by = c("hospitalization_id"))

```

# PART SIX: Upload Patient Assessments, Vitals
```{r}
#Vitals
specific_vital_categories <- c("height_cm", "weight_kg", "sbp", "dbp", "map", "spo2", "temp_c", "heart_rate", "respiratory_rate")
vitals_all <- as.data.table(read_data(paste0(data_dir, "clif_vitals", file_type), 
  select = c("hospitalization_id", "recorded_dttm", "vital_category", "vital_value")))[vital_category %in% specific_vital_categories & hospitalization_id %in% part_five_imv_runs$hospitalization_id]

#Patient Assessments
specific_assessment_categories <- c("RASS", "gcs_total")
patient_assessments <- as.data.table(read_data(paste0(data_dir, "clif_patient_assessments", file_type), select = c("hospitalization_id", "recorded_dttm", "assessment_category", "numerical_value", "categorical_value")))[assessment_category %in% specific_assessment_categories & hospitalization_id %in% part_five_imv_runs$hospitalization_id]
patient_assessments[, `:=`(hospitalization_id = as.character(hospitalization_id))]
patient_assessments <- unique(patient_assessments, by = c("hospitalization_id", "recorded_dttm", "assessment_category"))

```

# PART SEVEN: Height, Weight, and BMI Data
```{r}
# Create Hospitalization Table for Filtered Hospitalizations 
hospitalization_table_part_seven <- hospitalization[hospitalization_id %in% part_five_imv_runs$hospitalization_id]
setnames(hospitalization_table_part_seven, "admission_dttm", "cohort_admission_dttm")
setnames(hospitalization_table_part_seven, "discharge_dttm", "cohort_discharge_dttm")
setnames(hospitalization_table_part_seven, "hospitalization_id", "cohort_hospitalization_id")
hospitalization_table_part_seven[, cohort_admission_dttm := as.POSIXct(cohort_admission_dttm, tz = time_zone)]

#Create time_markers
hospitalization_table_part_seven[, `12_months_after` := add_with_rollback(cohort_admission_dttm, months(12))]
hospitalization_table_part_seven[, `12_months_before` := add_with_rollback(cohort_admission_dttm, months(-12))]

# Create Table of All Hospitalizations for Patients Filtered Above
hosp_height_weight_all <- hospitalization[patient_id %in% hospitalization_table_part_seven$patient_id]
hosp_height_weight_all <- hosp_height_weight_all[, .(hospitalization_id, patient_id, admission_dttm, discharge_dttm)]
setorder(hosp_height_weight_all, patient_id, admission_dttm)

#Merge Cohort and Height Weight Hospitalizations
hosp_height_weight_filt <- merge(hosp_height_weight_all, hospitalization_table_part_seven, by = "patient_id", allow.cartesian = TRUE)
hosp_height_weight_filt <- hosp_height_weight_filt %>%
  filter(admission_dttm >= `12_months_before` & admission_dttm <= `12_months_after`)
setDT(hosp_height_weight_filt)

set(hosp_height_weight_filt, j = "patient_cohort_id", 
    value = paste(hosp_height_weight_filt$patient_id, 
                  hosp_height_weight_filt$cohort_hospitalization_id, sep = "_"))

# Create Vitals Table 
height_weight <- c("height_cm", "weight_kg")
vitals_bmi <- vitals_all[hospitalization_id %in% hosp_height_weight_filt$hospitalization_id & vital_category %in% height_weight]

# Reshape the data to put height/weight into separate columns, using mean to handle duplicates
vitals_bmi_wide <- dcast(vitals_bmi, 
                   hospitalization_id + recorded_dttm ~ vital_category, 
                   value.var = "vital_value", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
vitals_bmi_wide <- vitals_bmi_wide[, .(hospitalization_id, recorded_dttm, weight_kg, height_cm)]
setorder(vitals_bmi_wide, hospitalization_id, recorded_dttm)

# Obtain Patient_IDs associated with BMI Hospitalizations and merge with reshaped table
pat_hosp_list <- hosp_height_weight_filt[, .(patient_id, hospitalization_id, patient_cohort_id, cohort_admission_dttm, cohort_hospitalization_id)]
vitals_bmi_patient <- merge(vitals_bmi_wide, pat_hosp_list, by = "hospitalization_id", all = TRUE)
setorder(vitals_bmi_patient, hospitalization_id, recorded_dttm)

# Table of All Height Values - filter out very low values
height_patient <- vitals_bmi_patient[!is.na(height_cm)]
height_patient <- height_patient[, .(patient_cohort_id, hospitalization_id, cohort_hospitalization_id, cohort_admission_dttm, recorded_dttm, height_cm)]
height_patient <- height_patient[height_cm >= 100]

# Find Median Height for each Patient within +/- 1 year of IMV hospitalization
median_height_per_patient <- height_patient[, .(
  median_height = median(height_cm, na.rm = TRUE)
), by = patient_cohort_id]

height_hosp <- merge(median_height_per_patient, hosp_height_weight_filt, by = "patient_cohort_id", all.x = TRUE)
height_hosp <- height_hosp[, .(hospitalization_id, patient_cohort_id, cohort_hospitalization_id, median_height)]
height_hosp_filtered <- height_hosp[
  hospitalization_id == cohort_hospitalization_id
]

# Create Weight Table and filter out improbably low weights
weight_patient <- vitals_bmi_patient[!is.na(weight_kg)]
weight_patient <- weight_patient[, .(patient_cohort_id, hospitalization_id, cohort_hospitalization_id, cohort_admission_dttm, recorded_dttm, weight_kg)]
weight_patient <- weight_patient[weight_kg >= 35]

# Calculate absolute time difference between admission_dttm and recorded_dttm
weight_patient[, time_diff := abs(difftime(cohort_admission_dttm, recorded_dttm, units = "secs"))]

# For each patient_id and admission_dttm, find the row with the minimum time_diff
admission_weight_table <- weight_patient[, .SD[which.min(time_diff)], by = .(patient_cohort_id, cohort_admission_dttm), .SDcols = c("cohort_admission_dttm", "weight_kg")]

# Create the final table with patient_id, admission_dttm, and admission_weight
final_weight_table <- admission_weight_table[, .(patient_cohort_id, weight_kg)]

# Create BMI Table, heigh_m column, and calculate BMI
patient_bmi_data <- merge(final_weight_table, height_hosp_filtered, by = c("patient_cohort_id"), 
  all = TRUE)
patient_bmi_data[, height_m := median_height / 100]
patient_bmi_data[, bmi := weight_kg / (height_m^2)]

# Cleaned BMI Table
patient_bmi_data_clean <- patient_bmi_data[, .(hospitalization_id, weight_kg, median_height, bmi)]
setnames(patient_bmi_data_clean, "median_height", "admission_height")

# BMI/Height/Weight Data for Hospitalizations Through Part Six
patient_bmi_data_final <- patient_bmi_data_clean[hospitalization_id %in% part_five_imv_runs$hospitalization_id]

# Filter out patients with no BMI data - removes ~111
patient_bmi_data_final <- patient_bmi_data_final[!is.na(bmi)]

# Update IMV Runs
xx_part_five_filt <- part_five_imv_runs[hospitalization_id %in% patient_bmi_data_final$hospitalization_id]

part_seven_imv_runs <- merge(xx_part_five_filt, patient_bmi_data_final, by = "hospitalization_id", all = TRUE)
part_seven_imv_runs <- unique(part_seven_imv_runs, by = c("hospitalization_id"))
```

#PART EIGHT: Demographics Cleaning
```{r}
hospitalization_table_part_eight <- hospitalization[hospitalization_id %in% part_seven_imv_runs$hospitalization_id]
patient_language <- patient_demographics[patient_id %in% hospitalization_table_part_eight$patient_id]

# Step 1: Count and sort languages, excluding "English"
language_counts <- patient_language[
  !language_category %in% c("English", "Unknown or NA", "Spanish"), 
  .N, 
  by = language_category
][order(-N)]
fwrite(language_counts, file = paste0(final_data, institution, "_non_english_non_spanish_language_counts.csv"))

# Get the top 5 most common languages
top_languages <- language_counts[1:5, language_category]

# Add top_five flag to patient_language
patient_language[, top_five := language_category %in% top_languages]

patient_language[, top_five_non_eng_non_spanish := fifelse(
  top_five == TRUE, 
  "top_five_non_eng_non_spanish",
  fifelse(language_category == "English", 
          "English", 
          fifelse(language_category == "Spanish", 
                  "Spanish", 
                  NA_character_))
)]

patient_language[, .N, by = top_five_non_eng_non_spanish][order(-N)]

patient_language[, non_eng := ifelse(language_category == "English", "English", "Non_English")]

#Merge Back
demographics_final <- merge(hospitalization_table_part_eight, patient_language, by = "patient_id", all = TRUE)

#Social Vulnerability Metric
# Check if SVM_SCORE column already exists
if (!"SVM_SCORE" %in% names(demographics_final)) {
  # Load SVM Data
  svm_data <- readr::read_csv(
    "https://raw.githubusercontent.com/acortizmd/CLIF_Deep_Sedation_Language/main/code/SVM_ZIPCODE_SCORES_All%20States_All%20Counties_2025-07-03.csv"
  )
  setDT(svm_data)

  svm_score <- svm_data[, .(ZIPCODE, SVM_SCORE)]
  setnames(svm_score, "ZIPCODE", "zipcode_five_digit")

  demographics_final <- merge(
    demographics_final,
    svm_score[, .(zipcode_five_digit, SVM_SCORE)],
    by = "zipcode_five_digit",
    all.x = TRUE
  )
}

#Clean Out Unknown Languages, Missing Demographic Info, and People Under 18 or over 120
demographics_clean <- demographics_final[
  language_category != "Unknown or NA" &
  language_category != "Sign Language" &
  !is.na(SVM_SCORE) & SVM_SCORE != "" &
  !is.na(sex_category) & sex_category != "" &
  !is.na(race_category) & race_category != "" &
  !is.na(ethnicity_category) & ethnicity_category != "" &
  age_at_admission >= 18 & age_at_admission < 120
]

demographics_clean_final <- demographics_clean[, .(hospitalization_id, patient_id, admission_dttm, discharge_dttm, age_at_admission, admission_type_category, discharge_category, sex_category, race_category, ethnicity_category, death_dttm, language_category, top_five_non_eng_non_spanish, non_eng, SVM_SCORE)]

demographics_clean_final[, race_category := as.character(race_category)]
demographics_clean_final[, race_category := fifelse(
  race_category %in% c("American Indian or Alaska Native", "Native Hawaiian or Other Pacific Islander", "Unknown", "Other"),
  "Other/Unknown",
  race_category
)]
demographics_clean_final[, ethnicity_category_final := fifelse(
  ethnicity_category %in% c("Non-Hispanic", "Unknown"),
  "Non-Hispanic/Unknown",
  ethnicity_category
)]

# Updated IMV
xx_part_seven_filt <- part_seven_imv_runs[hospitalization_id %in% demographics_clean_final$hospitalization_id]

part_eight_imv_runs <- merge(xx_part_seven_filt, demographics_clean_final, by = "hospitalization_id", all = TRUE)
part_eight_imv_runs[, death_dttm := as.POSIXct(death_dttm, tz = time_zone)]
part_eight_imv_runs[, begin_imv := as.POSIXct(begin_imv, tz = time_zone)]
```

# PART NINE: Elixhauser Comorbidity Index
```{r}
#Upload Hospital Diagnosis
hospital_diagnosis <- as.data.table(read_data(paste0(data_dir,"clif_hospital_diagnosis", file_type), select = c("hospitalization_id", "diagnostic_code", "diagnostic_code_format")))
setorder(hospital_diagnosis, hospitalization_id)

hospital_diagnosis_cohort <- hospital_diagnosis[hospitalization_id %in% part_eight_imv_runs$hospitalization_id]

lookback <- hospital_diagnosis_cohort %>%
  mutate(diagnostic_code = str_replace_all(diagnostic_code, "[^a-zA-Z0-9]", ""))

elixhauser_flags <- lookback %>%
  mutate(
    chf = str_detect(diagnostic_code, "^I099|^I110|^I130|^I132|^I255|^I420|^I425|^I426|^I427|^I428|^I429|^I43|^I50|^P290"),
    carit = str_detect(diagnostic_code, "^I441|^I442|^I443|^I456|^I459|^I47|^I48|^I49|^R000|^R001|^R008|^T821|^Z450|^Z950"),
    valv = str_detect(diagnostic_code, "^A520|^I05|^I06|^I07|^I08|^I091|^I098|^I34|^I35|^I36|^I37|^I38|^I39|^Q230|^Q231|^Q232|^Q233|^Z952|^Z953|^Z954"),
    pcd = str_detect(diagnostic_code, "^I26|^I27|^I280|^I288|^I289"),
    pvd = str_detect(diagnostic_code, "^I70|^I71|^I731|^I738|^I739|^I771|^I790|^I792|^K551|^K558|^K559|^Z958|^Z959"),
    hypunc = str_detect(diagnostic_code, "^I10"),
    hypc = str_detect(diagnostic_code, "^I11|^I12|^I13|^I15"),
    para = str_detect(diagnostic_code, "^G041|^G114|^G801|^G802|^G81|^G82|^G830|^G831|^G832|^G833|^G834|^G839"),
    ond = str_detect(diagnostic_code, "^G10|^G11|^G12|^G13|^G20|^G21|^G22|^G254|^G255|^G312|^G318|^G319|^G32|^G35|^G36|^G37|^G40|^G41|^G931|^G934|^R470|^R56"),
    cpd = str_detect(diagnostic_code, "^I278|^I279|^J40|^J41|^J42|^J43|^J44|^J45|^J46|^J47|^J60|^J61|^J62|^J63|^J64|^J65|^J66|^J67|^J684|^J701|^J703"),
    diabunc = str_detect(diagnostic_code, "^E10[0-19]|^E11[0-19]|^E12[0-29]|^E13[0-39]|^E14[0-49]$"),
    diabc = str_detect(diagnostic_code, "^E10[2-8]|^E11[2-8]|^E12[2-8]|^E13[2-8]|^E14[2-8]"),
    hypothy = str_detect(diagnostic_code, "^E00|^E01|^E02|^E03|^E890"),
    rf = str_detect(diagnostic_code, "^I120|^I131|^N18|^N19|^N250|^Z490|^Z491|^Z492|^Z940|^Z992"),
    ld = str_detect(diagnostic_code, "^B18|^I85|^I864|^I982|^K70|^K711|^K713|^K714|^K715|^K717|^K72|^K73|^K74|^K760|^K762|^K763|^K764|^K765|^K766|^K767|^K768|^K769|^Z944"),
    pud = str_detect(diagnostic_code, "^K257|^K259|^K267|^K269|^K277|^K279|^K287|^K289"),
    aids = str_detect(diagnostic_code, "^B20|^B21|^B22|^B24"),
    lymph = str_detect(diagnostic_code, "^C81|^C82|^C83|^C84|^C85|^C88|^C96|^C900|^C902"),
    metacanc = str_detect(diagnostic_code, "^C77|^C78|^C79|^C80"),
    solidtum = str_detect(diagnostic_code, "^C0[0-9]|^C1[0-4]|^C15|^C16|^C17|^C18|^C19|^C20|^C21|^C22|^C23|^C24|^C25|^C26|^C30|^C31|^C32|^C33|^C34|^C37|^C38|^C39|^C40|^C41|^C43|^C45|^C46|^C47|^C48|^C49|^C50|^C51|^C52|^C53|^C54|^C55|^C56|^C57|^C58|^C60|^C61|^C62|^C63|^C64|^C65|^C66|^C67|^C68|^C69|^C70|^C71|^C72|^C73|^C74|^C75|^C76|^C97"),
    rheumd = str_detect(diagnostic_code, "^L940|^L941|^L943|^M05|^M06|^M08|^M120|^M123|^M30|^M310|^M311|^M312|^M313|^M32|^M33|^M34|^M35|^M45|^M461|^M468|^M469"),
    coag = str_detect(diagnostic_code, "^D65|^D66|^D67|^D68|^D691|^D693|^D694|^D695|^D696"),
    obes = str_detect(diagnostic_code, "^E66"),
    wloss = str_detect(diagnostic_code, "^E40|^E41|^E42|^E43|^E44|^E45|^E46|^R634|^R64"),
    fed = str_detect(diagnostic_code, "^E222|^E86|^E87"),
    blane = str_detect(diagnostic_code, "^D500"),
    dane = str_detect(diagnostic_code, "^D508|^D509|^D51|^D52|^D53"),
    alcohol = str_detect(diagnostic_code, "^F10|^E52|^G621|^I426|^K292|^K700|^K703|^K709|^T51|^Z502|^Z714|^Z721"),
    drug = str_detect(diagnostic_code, "^F1[1-6]|^F18|^F19|^Z715|^Z722"),
    psycho = str_detect(diagnostic_code, "^F20|^F22|^F23|^F24|^F25|^F28|^F29|^F302|^F312|^F315"),
    depre = str_detect(diagnostic_code, "^F204|^F313|^F314|^F315|^F32|^F33|^F341|^F412|^F432")
  ) %>%
  group_by(hospitalization_id) %>%
  summarise(across(everything(), ~ max(.x, na.rm = TRUE)))

elixhauser <- elixhauser_flags %>%
  mutate(elixhauser_score =
           7 * chf + 5 * carit - 1 * valv + 4 * pcd + 2 * pvd +
           7 * para + 6 * ond + 3 * cpd + 5 * rf + 11 * ld +
           9 * lymph + 12 * metacanc + 4 * solidtum + 3 * coag -
           4 * obes + 6 * wloss + 5 * fed - 2 * blane - 2 * dane -
           7 * drug - 3 * depre)
setDT(elixhauser)

elixhauser_final <- elixhauser[, .(hospitalization_id, elixhauser_score)]

part_nine_imv_runs <- merge(part_eight_imv_runs, elixhauser_final, by = "hospitalization_id", all = TRUE)

part_nine_imv_runs <- part_nine_imv_runs[
  !is.na(elixhauser_score) & elixhauser_score != ""
]

```

# PART TEN: Remove patients with RASS values >8 hours from Beginning IMV and those with less than 2 RASS measurements per 24-hour period
```{r}
# RASS Assessment Table
pat_assess_rass <- patient_assessments[assessment_category == "RASS"]
setorder(pat_assess_rass, hospitalization_id, recorded_dttm)

pat_assess_rass_filt <- pat_assess_rass[hospitalization_id %in% part_nine_imv_runs$hospitalization_id]

# Convert Numerical and Categorical RASS Values to Standard Number
pat_assess_rass_filt[, rass_value := fifelse(
  !is.na(numerical_value), # If numerical_value is present, prioritize it
  as.character(as.integer(numerical_value)), # Convert numerical_value to integer and then to character
  fifelse(
    grepl("^\\+?(-?5|-?4|-?3|-?2|-?1|0|1|2|3|4)$", categorical_value), # Check if categorical_value matches permissible values (with optional + or -)
    as.character(as.integer(sub("^\\+", "", categorical_value))), # Remove leading "+" if present, then convert to integer and then to character
    NA_character_ ))]

# Merge imv_runs from part 8 with RASS filtered
hosp_time_rass <- merge(part_nine_imv_runs, pat_assess_rass_filt, by = "hospitalization_id", all = TRUE)

# Calculate Time Difference Between RASS Documentation and Begin IMV
hosp_time_rass[, time_to_first_rass := as.numeric(difftime(recorded_dttm, begin_imv, units = "hours"))]

# Determine RASS value closest to Beginning IMV
hosp_time_rass_closest <- hosp_time_rass[time_to_first_rass >= -1, 
                                         .SD[which.min(time_to_first_rass)], 
                                         by = hospitalization_id]

# Clean Out Those with RASS Values Not Documented within 8 hours of IMV
# Removes ~900 Admissions
time_to_first_rass_final <- hosp_time_rass_closest[time_to_first_rass >= -1 & time_to_first_rass <= 8]
time_to_first_rass_final[time_to_first_rass < 0, time_to_first_rass := 0]
setnames(time_to_first_rass_final, "recorded_dttm", "first_rass_dttm")
time_to_first_rass_final <- time_to_first_rass_final[, .(hospitalization_id, time_to_first_rass)]


# Timestamps
rass_cohort_timestamps <- part_nine_imv_runs[hospitalization_id %in% time_to_first_rass_final$hospitalization_id][, .(hospitalization_id, begin_imv, end_imv, imv_duration_hours)]

# Calculate start_time (begin_imv)
rass_cohort_timestamps[, start_time := as.POSIXct(begin_imv), by = hospitalization_id]

# Calculate end_time based on the condition for each hospitalization_id
rass_cohort_timestamps[, end_time := fifelse(
  imv_duration_hours >= 48,
  as.POSIXct(begin_imv) + hours(48),
  as.POSIXct(end_imv)
), by = hospitalization_id]

pat_assess_rass_cohort_clean <- pat_assess_rass_filt[hospitalization_id %in% time_to_first_rass_final$hospitalization_id]
setorder(pat_assess_rass_cohort_clean, hospitalization_id, recorded_dttm)

#Clean Only for rass_value
total_timeline_rass <- merge(pat_assess_rass_cohort_clean, rass_cohort_timestamps, by = "hospitalization_id", all = TRUE)[, .(hospitalization_id, recorded_dttm, rass_value, start_time, end_time)]

# Filter the table to retain rows where recorded_dttm is within the time range
total_timeline_rass_filt <- total_timeline_rass[recorded_dttm >= start_time & recorded_dttm <= end_time]

total_timeline_rass_filt <- total_timeline_rass_filt[!is.na(rass_value)]
total_timeline_rass_filt[, rass_value := as.numeric(rass_value)]

# Ensure data is sorted chronologically within each hospitalization
setorder(total_timeline_rass_filt, hospitalization_id, recorded_dttm)

##Median RASS Values
rass_median_by_hosp <- total_timeline_rass_filt[, .(
  median_rass_cont = median(rass_value, na.rm = TRUE)
), by = hospitalization_id]
##

# Compute the time duration for which each RASS value persists
total_timeline_rass_filt[, lead_time := shift(recorded_dttm, type = "lead"), by = hospitalization_id]

# If it's the last recorded RASS value, set duration to the difference between end_time and recorded_dttm
total_timeline_rass_filt[, rass_duration := as.numeric(difftime(lead_time, recorded_dttm, units = "hours"))]
total_timeline_rass_filt[is.na(rass_duration), rass_duration := as.numeric(difftime(end_time, recorded_dttm, units = "hours"))]

# Calculate the time-weighted sum of RASS values per hospitalization
rass_weighted_sum <- total_timeline_rass_filt[, .(rass_weighted = sum(rass_value * rass_duration, na.rm = TRUE)), by = hospitalization_id]

# Calculate total time spent under observation
rass_total_time <- total_timeline_rass_filt[, .(total_observed_time = sum(rass_duration, na.rm = TRUE)), by = hospitalization_id]

# Merge to get the final time-weighted RASS average
rass_calc <- merge(rass_weighted_sum, rass_total_time, by = "hospitalization_id", all = TRUE)
rass_calc[, rass_weighted_avg := rass_weighted / total_observed_time]

# Define deep sedation based on the time-weighted RASS average
rass_calc[, deep_sedation_weighted := rass_weighted_avg <= -3]

# Merge with timestamps to get full dataset
rass_calc_timeline <- merge(rass_calc, rass_cohort_timestamps, by = "hospitalization_id", all = TRUE)

# Set observation time (capped at 48 hours)
rass_calc_timeline[, observation_time := ifelse(imv_duration_hours < 48, imv_duration_hours, 48)]

rass_measurements <- total_timeline_rass_filt[, .(rass_count = .N), by = hospitalization_id]
rass_cumulative <- total_timeline_rass_filt[, .(rass_cumulative = sum(rass_value, na.rm = TRUE)), 
                                            by = hospitalization_id]
rass_cumulative_neg <- total_timeline_rass_filt[, .(
  rass_cumulative_neg = sum(ifelse(rass_value < 0, rass_value, 0), na.rm = TRUE)
), by = hospitalization_id]

rass_calc_timeline_merge <- merge(rass_calc_timeline, rass_measurements, by = "hospitalization_id", all = TRUE)
rass_calc_timeline_merge_two <- merge(rass_calc_timeline_merge, rass_cumulative, by = "hospitalization_id", all = TRUE)
rass_calc_timeline_merge_three <- merge(rass_calc_timeline_merge_two, rass_cumulative_neg, by = "hospitalization_id", all = TRUE)
rass_calc_timeline_merge_four <- merge(rass_calc_timeline_merge_three, rass_median_by_hosp, by = "hospitalization_id", all = TRUE)

# Compute RASS frequency (number of RASS values per observation period)
rass_calc_timeline_merge_four[, rass_frequency_hrs := observation_time / rass_count]
 
rass_calc_timeline_merge_four[, rass_unweighted_avg := rass_cumulative / rass_count]

# Define deep sedation based on the time-weighted RASS average
rass_calc_timeline_merge_four[, deep_sedation_unweighted := rass_unweighted_avg < -3]

# Define deep sedation based on median RASS
rass_calc_timeline_merge_four[, median_rass_binary := median_rass_cont <= -3]

# Sedation Index Score
rass_calc_timeline_merge_four[, sedation_index_score := rass_cumulative_neg / rass_count]

# Total Number of RASS Values -3 or lower
rass_counts <- total_timeline_rass_filt[rass_value <= -3, .N, by = hospitalization_id]
setnames(rass_counts, "N", "rass_below_minus_3_count")

rass_calc_timeline_merge_five <- merge(rass_calc_timeline_merge_four, rass_counts, by = "hospitalization_id", all = TRUE)

# Apply filtering
rass_calc_timeline_filt <- rass_calc_timeline_merge_five[rass_frequency_hrs <= 12][, .(hospitalization_id, rass_weighted_avg, rass_unweighted_avg, rass_frequency_hrs, rass_count, deep_sedation_weighted, deep_sedation_unweighted, rass_below_minus_3_count, median_rass_binary, median_rass_cont, sedation_index_score)]
setnafill(rass_calc_timeline_filt, cols = "rass_below_minus_3_count", fill = 0)


#GCS Processing

# GCS Assessment Table
pat_assess_gcs_proc <- patient_assessments[assessment_category == "gcs_total"]
setorder(pat_assess_gcs_proc, hospitalization_id, recorded_dttm)

pat_assess_gcs_filt <- pat_assess_gcs_proc[hospitalization_id %in% rass_calc_timeline_filt$hospitalization_id]

rass_cohort_timestamps_filt <- rass_cohort_timestamps[hospitalization_id %in% rass_calc_timeline_filt$hospitalization_id]

#Clean Only for gcs_value
total_timeline_gcs <- merge(pat_assess_gcs_filt, rass_cohort_timestamps_filt, by = "hospitalization_id", all = TRUE)[, .(hospitalization_id, recorded_dttm, numerical_value, start_time, end_time)]

# Filter the table to retain rows where recorded_dttm is within the time range
total_timeline_gcs <- total_timeline_gcs[recorded_dttm >= start_time & recorded_dttm <= end_time]

total_timeline_gcs <- total_timeline_gcs[!is.na(numerical_value)]
total_timeline_gcs[, numerical_value := as.numeric(numerical_value)]

# Ensure data is sorted chronologically within each hospitalization
setorder(total_timeline_gcs, hospitalization_id, recorded_dttm)

##Median GCS Values
gcs_median_by_hosp <- total_timeline_gcs[, .(
  median_gcs = median(numerical_value, na.rm = TRUE)
), by = hospitalization_id]
##

##Mean GCS Values
gcs_mean_by_hosp <- total_timeline_gcs[, .(
  mean_gcs = mean(numerical_value, na.rm = TRUE)
), by = hospitalization_id]
##

#Merge GCS
gcs_values <- merge(gcs_median_by_hosp, gcs_mean_by_hosp, by = "hospitalization_id", all = TRUE)

#Merge GCS and RASS

rass_calc_timeline_filt_final <- merge(rass_calc_timeline_filt, gcs_values, by = "hospitalization_id", all = TRUE)

# Updated IMV
xx_part_nine_filt <- part_nine_imv_runs[hospitalization_id %in% rass_calc_timeline_filt_final$hospitalization_id]

part_ten_imv_runs <- merge(xx_part_nine_filt, rass_calc_timeline_filt_final, by = "hospitalization_id", all = TRUE)

#Choose shortest imv_duration
part_ten_imv_runs[, imv_duration_hours_final := 
  fifelse(
    is.na(imv_duration_hours) & is.na(imv_duration_hours_trach), 
    NA_real_,
    pmin(imv_duration_hours, imv_duration_hours_trach, na.rm = TRUE)
  )
]
part_ten_imv_runs[imv_duration_hours_final>= 0, ] #JHU Patch
part_ten_imv_runs <- unique(part_ten_imv_runs, by = c("hospitalization_id"))
```

#PART ELEVEN: Assemblance of Time-Independent Covariate Table
```{r}
# Correct any death_dttm errors
part_ten_imv_runs[death_dttm < discharge_dttm & discharge_category == "Expired", death_dttm := discharge_dttm]
part_ten_imv_runs[
  is.na(death_dttm) & discharge_category == "Expired", 
  death_dttm := discharge_dttm
]

# Died During Admission Flag
part_ten_imv_runs[, death_during_admission := 
    !is.na(death_dttm) & (death_dttm >= admission_dttm & death_dttm <= discharge_dttm)]

# Total IMV Time in Days
part_ten_imv_runs[, imv_duration_days := imv_duration_hours_final / 24]

# Died Within 28-Days
part_ten_imv_runs[, death_within_28_days_of_imv := 
    !is.na(death_dttm) & !is.na(begin_imv) & 
    (death_dttm >= begin_imv - days(1) & death_dttm <= begin_imv + days(28))]

# Days from Begin IMV to Death
part_ten_imv_runs[, days_from_begin_imv_to_death := 
    ifelse(death_within_28_days_of_imv == TRUE & abs(as.numeric(difftime(death_dttm, begin_imv, units = "days")) - imv_duration_days) <= 1, 
           imv_duration_days, 
           ifelse(death_within_28_days_of_imv == TRUE, 
                  as.numeric(difftime(death_dttm, begin_imv, units = "days")), 
                  NA))]

# 30 Day Mortality
part_ten_imv_runs[, thirty_day_mortality := 
    !is.na(death_dttm) & !is.na(begin_imv) & 
    (death_dttm >= begin_imv - days(1) & death_dttm <= begin_imv + days(30))]

#Age Category
part_ten_imv_runs <- part_ten_imv_runs %>%
  mutate(age_category = case_when(
    age_at_admission >= 18 & age_at_admission < 35 ~ "18-34 Years",
    age_at_admission >= 35 & age_at_admission < 45 ~ "35-44 Years",
    age_at_admission >= 45 & age_at_admission < 55 ~ "45-54 Years",
    age_at_admission >= 55 & age_at_admission < 65 ~ "55-64 Years",
    age_at_admission >= 65 & age_at_admission < 75 ~ "65-74 Years",
    age_at_admission >= 75  ~ "75+ Years",
    TRUE ~ NA_character_ # Assign NA for any cases that don't fit
  ))
setDT(part_ten_imv_runs)

#############################################
#Hospital Free Days over 6 months

#Table of hospitalization_id & admission_dttm for all encounters
cohort_admissions <- part_ten_imv_runs[, .(hospitalization_id, patient_id, admission_dttm, discharge_dttm)]
setnames(cohort_admissions, "admission_dttm", "cohort_admission_dttm")
setnames(cohort_admissions, "discharge_dttm", "cohort_discharge_dttm")
setnames(cohort_admissions, "hospitalization_id", "cohort_hospitalization_id")
cohort_admissions[, cohort_admission_dttm := as.POSIXct(cohort_admission_dttm, tz = time_zone)]

#Create 6_month_marker
cohort_admissions[, `6_month` := add_with_rollback(cohort_admission_dttm, months(6))]

#All hospitalizations for the cohort
hosp_hfd <- hospitalization[patient_id %in% part_ten_imv_runs$patient_id]

#Merge with cohort_admissions
hfd_calc <- merge(hosp_hfd, cohort_admissions, by = "patient_id", allow.cartesian = TRUE)

#Filter hosp_hfd only for those between cohort_admission_dttm and 6 months of cohort_admission_dttm
hfd_calc_filtered <- hfd_calc %>%
  filter(admission_dttm >= cohort_admission_dttm & admission_dttm <= `6_month`)

#Calculate Total Hospitalization Days for each patient_id taking into account 6-month cutoff
hfd_calc_filtered <- hfd_calc_filtered %>%
  mutate(total_days = ifelse(
    discharge_dttm <= `6_month`, 
    as.numeric(difftime(discharge_dttm, admission_dttm, units = "days")),
    as.numeric(difftime(`6_month`, admission_dttm, units = "days"))
  ))
setDT(hfd_calc_filtered)

set(hfd_calc_filtered, j = "patient_cohort_id", 
    value = paste(hfd_calc_filtered$patient_id, 
                  hfd_calc_filtered$cohort_hospitalization_id, sep = "_"))

hfd_calc_final <- hfd_calc_filtered %>%
  group_by(patient_cohort_id) %>%
  mutate(hospitalization_days_6_month_total = sum(total_days, na.rm = TRUE)) %>%
  ungroup()
setDT(hfd_calc_final)
setorder(hfd_calc_final, patient_cohort_id, admission_dttm)

hfd_calc_final_filtered <- hfd_calc_final[
  hospitalization_id == cohort_hospitalization_id
]

hfd_calc_final_filtered <- hfd_calc_final_filtered[, .(hospitalization_id, hospitalization_days_6_month_total)]

part_ten_imv_runs_intermediate <- merge(part_ten_imv_runs, hfd_calc_final_filtered, by = "hospitalization_id", all = TRUE)

# Died Within 6-Months
part_ten_imv_runs_intermediate[, death_within_6_months_of_admission := 
    !is.na(death_dttm) & !is.na(admission_dttm) & 
    (death_dttm >= admission_dttm - days(1) & death_dttm <= admission_dttm + days(183))]

# Days from Admission to Death
part_ten_imv_runs_intermediate[, days_from_admission_to_death := 
           ifelse(death_within_6_months_of_admission == TRUE, 
                  as.numeric(difftime(death_dttm, admission_dttm, units = "days")), 
                  NA)]

# Six Month Hospital Free Days
part_ten_imv_runs_intermediate[, hfd_6m := 
    ifelse(hospitalization_days_6_month_total > 183, 
           0,  # Case 0: More than 183 hospitalization days
           ifelse(death_during_admission == TRUE, 
                  0,  # Case 1: Death during admission
                  ifelse(death_within_6_months_of_admission == FALSE, 
                         pmax(183 - hospitalization_days_6_month_total, 0),  # Case 2: Survived 6 months, ensure non-negative
                         ifelse(death_within_6_months_of_admission == TRUE & death_during_admission == FALSE, 
                                pmax(days_from_admission_to_death - hospitalization_days_6_month_total, 0),  # Case 3: Died within 6 months but not during admission, ensure non-negative
                                NA  # Catch-all for unexpected cases
                         )
                  )
           )
    )
]

#################################
# 28 Day Hospital Free Days

#Create 6_month_marker
cohort_admissions[, `28_days` := add_with_rollback(cohort_admission_dttm, days(28))]

#All hospitalizations for the cohort
hosp_hfd_28 <- hospitalization[patient_id %in% part_ten_imv_runs$patient_id]

#Merge with cohort_admissions
hfd_calc_28 <- merge(hosp_hfd_28, cohort_admissions, by = "patient_id", allow.cartesian = TRUE)

#Filter hosp_hfd only for those between cohort_admission_dttm and 6 months of cohort_admission_dttm
hfd_calc_filtered_28 <- hfd_calc_28 %>%
  filter(admission_dttm >= cohort_admission_dttm & admission_dttm <= `28_days`)

#Calculate Total Hospitalization Days for each patient_id taking into account 6-month cutoff
hfd_calc_filtered_28 <- hfd_calc_filtered_28 %>%
  mutate(total_days = ifelse(
    discharge_dttm <= `28_days`, 
    as.numeric(difftime(discharge_dttm, admission_dttm, units = "days")),
    as.numeric(difftime(`28_days`, admission_dttm, units = "days"))
  ))
setDT(hfd_calc_filtered_28)

set(hfd_calc_filtered_28, j = "patient_cohort_id", 
    value = paste(hfd_calc_filtered_28$patient_id, 
                  hfd_calc_filtered_28$cohort_hospitalization_id, sep = "_"))

hfd_calc_final_28 <- hfd_calc_filtered_28 %>%
  group_by(patient_cohort_id) %>%
  mutate(hospitalization_days_28_days_total = sum(total_days, na.rm = TRUE)) %>%
  ungroup()
setDT(hfd_calc_final_28)
setorder(hfd_calc_final_28, patient_cohort_id, admission_dttm)

hfd_calc_final_filtered_28 <- hfd_calc_final_28[
  hospitalization_id == cohort_hospitalization_id
]

hfd_calc_final_filtered_28 <- hfd_calc_final_filtered_28[, .(hospitalization_id, hospitalization_days_28_days_total)]

part_ten_imv_runs_3 <- merge(part_ten_imv_runs_intermediate, hfd_calc_final_filtered_28, by = "hospitalization_id", all = TRUE)

# Died Within 6-Months
part_ten_imv_runs_3[, death_within_28_days_of_admission := 
    !is.na(death_dttm) & !is.na(admission_dttm) & 
    (death_dttm >= admission_dttm - days(1) & death_dttm <= admission_dttm + days(28))]

# Days from Admission to Death
part_ten_imv_runs_3[, days_from_admission_to_death := 
           ifelse(death_within_28_days_of_admission == TRUE, 
                  as.numeric(difftime(death_dttm, admission_dttm, units = "days")), 
                  NA)]

# Six Month Hospital Free Days
part_ten_imv_runs_3[, hfd_28d := 
    ifelse(hospitalization_days_28_days_total > 28, 
           0,  # Case 0: More than 28 hospitalization days
           ifelse(death_during_admission == TRUE, 
                  0,  # Case 1: Death during admission
                  ifelse(death_within_28_days_of_admission == FALSE, 
                         pmax(28 - hospitalization_days_28_days_total, 0),  # Case 2: Survived 28 days, ensure non-negative
                         ifelse(death_within_28_days_of_admission == TRUE & death_during_admission == FALSE, 
                                pmax(days_from_admission_to_death - hospitalization_days_28_days_total, 0),  # Case 3: Died within 6 months but not during admission, ensure non-negative
                                NA  # Catch-all for unexpected cases
                         )
                  )
           )
    )
]


###################################
# 28 Day Ventilator Free Days
#First Deal with Reintubations
reintubations_vfd <- reintubations[hospitalization_id %in% part_ten_imv_runs$hospitalization_id]
reintubations_vfd <- reintubations_vfd[, .(hospitalization_id, begin_imv, end_imv, run_id)]

begin_first_imv <- part_ten_imv_runs[hospitalization_id %in% reintubations_vfd$hospitalization_id]
  
begin_first_imv <- begin_first_imv[, .(hospitalization_id, begin_imv)]
setnames(begin_first_imv, "begin_imv", "begin_first_imv")

vfd_reintubated_patients <- merge(reintubations_vfd, begin_first_imv, by = "hospitalization_id", all.x = TRUE)
vfd_reintubated_patients[, `28_day` := begin_first_imv + days(28)]
vfd_reintubated_patients[, `28_day_flag` := begin_imv < `28_day`]

vfd_reintubated_patients[, days_reintubated := 
  ifelse(`28_day_flag` == TRUE, 
         as.numeric(pmin(end_imv, `28_day`) - begin_imv, units = "days"),
         NA_real_)
]

days_reintubated <- vfd_reintubated_patients[
  , .(days_reintubated_28 = sum(days_reintubated, na.rm = TRUE)), 
  by = hospitalization_id
]

part_ten_imv_runs_final <- merge(part_ten_imv_runs_3, days_reintubated, by = "hospitalization_id", all = TRUE)
setnafill(part_ten_imv_runs_final, cols = "days_reintubated_28", fill = 0)

part_ten_imv_runs_final[, vfd_raw := 
    ifelse(death_within_28_days_of_imv == FALSE & imv_duration_days < 28, 
           28 - imv_duration_days, 
           ifelse(death_within_28_days_of_imv == FALSE & imv_duration_days > 28, 
                  0, 
                  ifelse(death_within_28_days_of_imv == TRUE, 
                         days_from_begin_imv_to_death - imv_duration_days, 
                         NA)))]

part_ten_imv_runs_final[, vfd_clean := vfd_raw - days_reintubated_28]
#################### NEED TO FIGURE THIS OUT ########################
###28 Day ICU Free Days

begin_imv <- part_ten_imv_runs_final[, .(patient_id, hospitalization_id, begin_imv)]

#Create 28_day_marker
begin_imv[, `28_day` := add_with_rollback(begin_imv, days(28))]
setnames(begin_imv, "hospitalization_id", "cohort_hospitalization_id")

#Merge with cohort_admissions
icufd_calc <- merge(hosp_hfd, begin_imv, by = c("patient_id"), allow.cartesian = TRUE)

#Filter hosp_hfd only for those between cohort_admission_dttm and 28 days of cohort_admission_dttm
icufd_calc_filtered <- icufd_calc %>%
  filter(admission_dttm >= begin_imv & admission_dttm <= `28_day`)
#These are all of the repeat hospitalizations

#Need to determine the amount of days before 28_day spent in ICU for each of these hospitalizations
adt_repeat_hosp <- adt[hospitalization_id %in% icufd_calc_filtered$hospitalization_id]

#Merge
icufd_adt_merge <- merge(adt_repeat_hosp, icufd_calc_filtered, by = "hospitalization_id", all = TRUE)

# Filter for ICU rows only
icu_only_repeat <- icufd_adt_merge[location_category == "icu"]

# Clip both ends of the interval
icu_only_repeat[, icu_start := pmax(in_dttm, begin_imv, na.rm = TRUE)]
icu_only_repeat[, icu_end := pmin(out_dttm, `28_day`, na.rm = TRUE)]

# Keep only valid intervals
icu_only_repeat <- icu_only_repeat[icu_start < icu_end]

# Calculate ICU time in hours
icu_only_repeat[, icu_hours := as.numeric(difftime(icu_end, icu_start, units = "hours"))]
icu_only_repeat_filt <- icu_only_repeat[hospitalization_id != cohort_hospitalization_id]

# Remove negative or NA durations just in case
icu_only_repeat_filt <- icu_only_repeat_filt[!is.na(icu_hours) & icu_hours > 0]

# Sum ICU hours per cohort_hospitalization_id
icu_time_summary_repeat <- icu_only_repeat_filt[, .(icu_hours_repeat_hosp = sum(icu_hours)), by = cohort_hospitalization_id]

setnames(icu_time_summary_repeat, "cohort_hospitalization_id", "hospitalization_id")

#NOW PROCESS THE OG HOSPITALIZATIONS
icufd_calc_og <- begin_imv
setnames(icufd_calc_og, "cohort_hospitalization_id", "hospitalization_id")
adt_og <- adt[hospitalization_id %in% begin_imv$hospitalization_id]

#Merge
icufd_adt_merge_og <- merge(adt_og, icufd_calc_og, by = "hospitalization_id", all = TRUE)

# Filter for ICU rows only
icu_only_og <- icufd_adt_merge_og[location_category == "icu"]

# Clip each out_dttm to the 28_day timestamp if it goes beyond it
icu_only_og[, icu_start := pmax(in_dttm, begin_imv, na.rm = TRUE)]
icu_only_og[, icu_end := pmin(out_dttm, `28_day`, na.rm = TRUE)]

icu_only_og <- icu_only_og[icu_start < icu_end]

icu_only_og[, icu_hours := as.numeric(difftime(icu_end, icu_start, units = "hours"))]

# Remove negative or NA durations just in case
icu_only_og <- icu_only_og[!is.na(icu_hours) & icu_hours > 0]

# Sum ICU hours per cohort_hospitalization_id
icu_time_summary_og <- icu_only_og[, .(icu_hours_og_hosp = sum(icu_hours)), by = hospitalization_id]

icu_time_summary_total <- merge(icu_time_summary_og, icu_time_summary_repeat, by = "hospitalization_id", all = TRUE)
setnafill(icu_time_summary_total, cols = "icu_hours_repeat_hosp", fill = 0)
icu_time_summary_total[, icu_hours_total := icu_hours_og_hosp + icu_hours_repeat_hosp]
icu_time_summary_total[, icu_days := icu_hours_total / 24]
icu_time_summary_total[, icufd_28 := 28 - icu_days]
icufd_final <- icu_time_summary_total[, .(hospitalization_id, icufd_28)]

part_eleven_imv_runs <- merge(part_ten_imv_runs_final, icufd_final, by = "hospitalization_id", all = TRUE)
```

# PART TWELVE: Upload CLIF Tables for Time-Dependent Covariates
```{r}
#Labs
specific_lab_categories <- c("creatinine", "bilirubin_total", "platelet_count", "po2_arterial", "pco2_arterial", "ph_arterial", "anion_gap", "troponin_t", "troponin_i", "sodium", "bicarbonate", "chloride", "bun", "albumin", "glucose_serum", "hemoglobin", "hematocrit", "lactate", "so2_arterial", "wbc")
labs_time_dependent <- as.data.table(read_data(paste0(data_dir,"clif_labs", file_type), select = c("hospitalization_id", "lab_order_dttm", "lab_collect_dttm", "lab_category", "lab_value_numeric", "lab_result_dttm")))[lab_category %in% specific_lab_categories]
setorder(labs_time_dependent, hospitalization_id, lab_collect_dttm)
setnames(labs_time_dependent, "lab_collect_dttm", "recorded_dttm")
labs_time_dependent_2 <- labs_time_dependent[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]

#Continuous Meds
specific_med_groups <- c("vasoactives", "sedation", "paralytics")
meds_all <- as.data.table(read_data(paste0(data_dir,"clif_medication_admin_continuous", file_type), select = c("hospitalization_id", "admin_dttm", "med_dose", "med_dose_unit", "med_category", "med_group")))[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id & med_group %in% specific_med_groups]
setorder(meds_all, hospitalization_id, admin_dttm)
setnames(meds_all, "admin_dttm", "recorded_dttm")

meds_time_dependent <- merge(part_eleven_imv_runs, meds_all, by = "hospitalization_id", all = TRUE)

meds_time_dependent <- meds_time_dependent %>%
  mutate(med_dose = case_when(
    !grepl("kg", med_dose_unit, ignore.case = TRUE) & !grepl("vasopressin|isoproterenol|fentanyl|morphine|hydromorphone|lorazepam|midazolam", med_category, ignore.case = TRUE) ~ med_dose / weight_kg,
    # Otherwise Keep the original med_dose
    TRUE ~ med_dose))
meds_time_dependent <- meds_time_dependent[!is.na(med_category)]
meds_time_dependent <- meds_time_dependent[, .(hospitalization_id, recorded_dttm, med_category, med_dose)]

#Meds Intermittent
intermittent_med_groups <- c("paralytics")
paralytic_pushes <- as.data.table(read_data(paste0(data_dir,"clif_medication_admin_intermittent", file_type), select = c("hospitalization_id", "admin_dttm", "med_dose", "med_category", "med_group")))[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id & med_group %in% intermittent_med_groups]
setorder(paralytic_pushes, hospitalization_id, admin_dttm)
#setnames(paralytic_pushes, "admin_dttm", "recorded_dttm")
reduced_hosp_info <- part_eleven_imv_runs[, .(hospitalization_id, begin_imv, end_imv)]
paralytics_hosp_merge <- merge(paralytic_pushes, reduced_hosp_info, by = "hospitalization_id", all = TRUE)
paralytics_hosp_merge <- paralytics_hosp_merge[!is.na(med_dose) & med_dose > 0]
paralytic_pushes_first_48h <- paralytics_hosp_merge[
  admin_dttm >= begin_imv + lubridate::hours(4) & admin_dttm <= begin_imv + lubridate::hours(48)
]
paralytic_pushes_first_48h[, push_paralytics := 1]
paralytic_pushes_first_48h_final <- paralytic_pushes_first_48h[, .(hospitalization_id, push_paralytics)]
paralytic_pushes_first_48h_final <- unique(paralytic_pushes_first_48h_final, by = c("hospitalization_id"))


#Filtered Vitals
time_dependent_vitals <- c("sbp", "dbp", "map", "spo2", "temp_c", "heart_rate", "respiratory_rate")
vitals_time_dependent <- vitals_all[vital_category %in% time_dependent_vitals & hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]

#Filtered Patient Assessments
pat_assess_rass_time_dependent <- pat_assess_rass_filt[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]
pat_assess_rass_time_dependent <- pat_assess_rass_time_dependent[, .(hospitalization_id, recorded_dttm, rass_value)]
pat_assess_rass_time_dependent[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]
pat_assess_rass_time_dependent[, rass_value := as.numeric(rass_value)]
patient_assess_gcs_time_dependent <- patient_assessments[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id & assessment_category == "gcs_total"]
pat_assess_gcs_time_dependent <- patient_assess_gcs_time_dependent[, .(hospitalization_id, recorded_dttm, assessment_category, numerical_value)]
#GCS
pat_assess_gcs_time_dependent_wide <- dcast(pat_assess_gcs_time_dependent, 
                   hospitalization_id + recorded_dttm ~ assessment_category, 
                   value.var = "numerical_value", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
setorder(pat_assess_gcs_time_dependent_wide, hospitalization_id, recorded_dttm)

pat_assess_gcs_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]
```

# LAPS2 Score Part 1: Encounter Data
```{r}
### change to hospital_icu
clif_adt_hospital <-  adt |>
  mutate(location_category = factor(location_category,
                                    levels = (c("icu", "stepdown", "ward", "ed", "procedural", "other")),                                    ordered = TRUE)) |> 
  group_by(hospitalization_id) |> 
  arrange(hospitalization_id, location_category, in_dttm) |> 
  
  # first one will be ICU
  slice_head(n=1) |> 
  ungroup() |> 
  dplyr::select(hospitalization_id,
         hospital = hospital_id)


clif_adt_admit <-  adt |>
  group_by(hospitalization_id) |> 
  arrange(hospitalization_id, in_dttm) |> 
  slice_head(n=1) |> 
  ungroup() |> 
  dplyr::select(hospitalization_id,
         first_location_category = location_category)
  

# fixing clif_labs to recorded_dttm
if ("lab_collect_dttm" %in% colnames(labs_time_dependent)) {
  labs_time_dependent <- labs_time_dependent |> rename(recorded_dttm = lab_collect_dttm)
}

df_encounter_laps2_temp1 <- hospitalization |> 
  left_join(patient_demographics) |> 
  left_join(clif_adt_hospital) |> 
  left_join(clif_adt_admit) |> 
  
  
  # make dead/hospice column
  mutate(
    death_or_hospice_01 = fcase(
      discharge_category %in% 
        c("hospice", "dead", "expired", "died", "Hospice", "Expired")       , 1,
      default                                         = 0
    ),
    
    death_or_hospice_01       = as.factor(death_or_hospice_01),
    
    # make 24 hours after admission variable (to collect data from first 24 hours)
    dt_24hours_after_admit    = admission_dttm + hours(24),
    
    # ED binary variable
    ed_admit_01               = fifelse(first_location_category == "ed", 1, 0, 0),
          # I think we used categorical in Rachel Kohn paper... Pat uses ED y/n
          # admission source OR you can use ED variable
    # Admit source variable (use one or the other)
    # admission_type_category      = as.factor(admission_type_category),  don't have this yet
    
    # sex_category
    female_01                 = fifelse(str_detect(sex_category, "Female"), 1, 0, 0),
    
  )
```

# LAPS2 Score Part 2:  Fix Anion Gap & Troponin_t
```{r}
if(any(labs_time_dependent$lab_category == "anion_gap", na.rm = TRUE)){
  
  print("you already have anion gap")
  message("we are going to remove anion_gap and recreate it for you")
  message("this way... all our analyses are the same")
  
  labs_time_dependent <- labs_time_dependent |> 
    filter(lab_category != "anion_gap")

    print("Anion gap removed... please proceed!!")
  
} else {
  
  print("no anion_gap found... we will make one for you!!")
  
  
}


if(any(labs_time_dependent$lab_category == "troponin_t")){
  
  print("you already have trop_t")
  message("we are going to change to trop_i levels")
  
  labs_time_dependent <- labs_time_dependent |> 
    
    mutate(lab_value_numeric = fcase(
      lab_category == "troponin_t",
      lab_value_numeric / 1000,
      default = lab_value_numeric
    )) |> 
    
    mutate(lab_category = fcase(
      lab_category == "troponin_t",
      "troponin_i",
      rep_len(TRUE, length(lab_category)), lab_category
    )) 
    
    print("Troponin fixed ... please proceed!!")
  
} else {
  
  print("no need to fix troponin !!")
  
}

# Set a time tolerance, e.g., 1 hour
time_tolerance <- hours(1)

# Create recorded_date from recorded_dttm
labs_time_dependent <- labs_time_dependent %>%
  mutate(recorded_date = as.Date(recorded_dttm))

# Separate the lab values for sodium, bicarbonate, and chloride, and create recorded_date
sodium_df <- labs_time_dependent %>%
  filter(lab_category == "sodium") %>%
  dplyr::select(hospitalization_id, recorded_dttm, recorded_date, sodium = lab_value_numeric,
         lab_order_dttm, lab_result_dttm)

bicarbonate_df <- labs_time_dependent %>%
  filter(lab_category == "bicarbonate") %>%
  dplyr::select(hospitalization_id, recorded_dttm, recorded_date, bicarbonate = lab_value_numeric)

chloride_df <- labs_time_dependent %>%
  filter(lab_category == "chloride") %>%
  dplyr::select(hospitalization_id, recorded_dttm, recorded_date, chloride = lab_value_numeric)

# Perform a join by clif_hospitalizations_joined_id and recorded_date to limit the join size
anion_gap_df <- sodium_df %>%
  full_join(bicarbonate_df, by = c("hospitalization_id", "recorded_date")) %>%
  full_join(chloride_df, by = c("hospitalization_id", "recorded_date")) %>%
  
  # Calculate the time differences between recorded_dttm values
  mutate(time_diff_sodium_bicarbonate = abs(difftime(recorded_dttm.x, recorded_dttm.y, units = "mins")),
         time_diff_sodium_chloride = abs(difftime(recorded_dttm.x, recorded_dttm, units = "mins")),
         
         # Apply the time tolerance (1 hour)
         within_tolerance_bicarbonate = time_diff_sodium_bicarbonate <= time_tolerance,
         within_tolerance_chloride = time_diff_sodium_chloride <= time_tolerance) %>%
  
  # Filter to retain rows where the time differences are within the tolerance
  filter(within_tolerance_bicarbonate & within_tolerance_chloride) %>%
  
  # Calculate the anion gap: sodium - (bicarbonate + chloride)
  mutate(anion_gap = sodium - (bicarbonate + chloride)) %>%
  
  # Select the relevant columns
  dplyr::select(hospitalization_id, recorded_dttm.x, anion_gap, lab_order_dttm, lab_result_dttm, recorded_date)

# Rename recorded_dttm.x to recorded_dttm for consistency
anion_gap_df <- anion_gap_df %>%
  mutate(reference_unit = "meq/l") |> 
  mutate(lab_name = "anion gap") |> 
  mutate(lab_category = "anion_gap") |>
  rename(lab_value_numeric = anion_gap) |> 
  rename(recorded_dttm = recorded_dttm.x)


labs_time_dependent <- labs_time_dependent |> 
              bind_rows(anion_gap_df)
```


# LAPS2 Score Part 3: Fix/Prep Labs for LAPS2 Risk Prediction (Prelaps)
```{r}
############
# Pre-Laps... want the most recent within 24 hours of admission
############

df_pre_laps2_prep1 <- 
  df_encounter_laps2_temp1 |> 
  left_join(labs_time_dependent |> 
              filter(lab_category %in% c("anion_gap",
                                         "bicarbonate", # FROM BMP
                                         "bun",
                                         "creatinine",
                                         "sodium")),
            
            # join only if between the first 24 hours
            by = join_by(hospitalization_id, 
                         dt_24hours_after_admit >= recorded_dttm) 
  ) |> 
  # Min/Max
  #     bun            max
  #     creatinine     min
  #     bun/cr         max
  #     bun/cr         most recent (prelaps)
  #     anion_gap      max
  #     bicarbonate min
  #     ag/bicarb      most recent (prelaps)
  #     sodium         min
  #     sodium         most recent (prelaps)
       
  group_by(hospitalization_id, recorded_dttm, lab_category) |> 
  # using obs lets you get away with duplicates when going wider and can fix later!
  mutate(obs = row_number()) |> 
  ungroup() |> 
  pivot_wider(
    id_cols = c(hospitalization_id, recorded_dttm, obs),
    names_from = lab_category,
    values_from = lab_value_numeric,
    # names_prefix = "lab_wide_",
    # values_fill = NA
  ) |>
  
  # Quick min/max for duplicates at the SAME time
  dplyr::select(-obs) |> 
  distinct() 

# find the duplicates by recorded_dttm
df_pre_laps2_prep2 <- df_pre_laps2_prep1 |> 
  arrange(hospitalization_id, recorded_dttm) |> 
  group_by(hospitalization_id, recorded_dttm) |>
  # mutate(n = n()) |> 
  # filter(n > 1) |> 
  summarise(
    bun               = fmax(bun, na.rm = TRUE),
    creatinine        = fmin(creatinine, na.rm = TRUE),
    bicarbonate       = fmin(bicarbonate, na.rm = TRUE),
    anion_gap         = fmax(anion_gap, na.rm = TRUE),
    sodium            = fmin(sodium, na.rm = TRUE),

    .groups = "drop"
  ) |> 
  
  # replace inf with NAs
  mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 
  
  # get rid of duplicate rows
  distinct() |> 
  
  # order appropriately for when getting last labs for "most recent"
  arrange(hospitalization_id, recorded_dttm) |> 
  
  # carry forward
  group_by(hospitalization_id) |> 
  
  # fill in ... its only the first 24 hours so its ok
  fill(c(bun,
         creatinine,
         bicarbonate,
         anion_gap,
         sodium), 
       .direction = "downup") |>
  
  # ? time shift if needed
  
  # get the most recent lab (aka last) based on recorded_dttm
    mutate(
      bun_recent                = flast(bun, na.rm = TRUE),
      creatinine_recent         = flast(creatinine, na.rm = TRUE),
      bicarbonate_recent        = flast(bicarbonate, na.rm = TRUE),
      anion_gap_recent          = flast(anion_gap, na.rm = TRUE),
      sodium_recent             = flast(sodium, na.rm = TRUE)
           ) |> 
  ungroup() |> 
  dplyr::select(hospitalization_id, ends_with("recent")) |> 

  # final pre-laps labs needed
  mutate(
    # make ratios
    sodium              = sodium_recent,
    bun_cr_ratio        = bun_recent/creatinine_recent,
    ag_hco3_ratio       = (anion_gap_recent/bicarbonate_recent)*1000
  ) |> 
  dplyr::select(-ends_with("recent")) |> 
  
  # replace inf with NAs (sometimes it happens when divided by zero)
  mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 
  
  distinct()
        
 df_pre_laps_merging <- df_pre_laps2_prep2 |> 
   left_join(
     df_encounter_laps2_temp1 |> 
       dplyr::select(
         hospitalization_id,
         admission_dttm,
         age = age_at_admission,
         female_01,
         ed_admit_01,
         # admission_type_category,
         death_or_hospice_01
       ) |> 
       distinct(),
     join_by(hospitalization_id)
   ) |> 
   mutate(
    age_cat = case_when(
      age >= 85 ~ 5,
      age >= 75 ~ 4,
      age >= 65 ~ 3,
      age >= 40 ~ 2,
      TRUE ~ 1
    )
  ) |> 
   mutate(
      # impute to normal (8)
    bun_cr_ratio = fifelse(is.na(bun_cr_ratio), 8, bun_cr_ratio),
     
    buncreat_cat = case_when(
      bun_cr_ratio        >= 24 ~ 4,
      bun_cr_ratio        >= 16 ~ 3,
      bun_cr_ratio        < 8   ~ 2,
      TRUE                      ~ 1
    )
  ) |> 
     mutate(
      # impute to normal (135)
    sodium = fifelse(is.na(sodium), 135, sodium),
    na_cat = case_when(
      sodium            >= 155  ~ 7,
      sodium            >= 149  ~ 6,
      sodium            >= 146  ~ 5,
      sodium            >= 132  ~ 4,
      sodium            >= 129  ~ 3,
      sodium            < 129   ~ 2,
      TRUE                      ~ 1
    )
  ) |> 
     mutate(
      # impute to normal (200)
    ag_hco3_ratio = fifelse(is.na(ag_hco3_ratio), 200, ag_hco3_ratio),
    aghco3_cat = case_when(
      ag_hco3_ratio        >= 600  ~ 4,
      ag_hco3_ratio        >= 400  ~ 3,
      ag_hco3_ratio        < 200   ~ 2,
      TRUE                         ~ 1,
    )
  )
 
 # logistic regression to obtain p(death/hospice)
 lr_model <-
  glm(
    death_or_hospice_01 ~
      age +
      female_01 +
      ed_admit_01 +
      bun_cr_ratio +
      sodium +
      ag_hco3_ratio,
    family = "binomial",
    data   = df_pre_laps_merging
  )

tidy(lr_model)
pR2(lr_model)["McFadden"]

df_pre_laps_final  <- df_pre_laps_merging |> 
  mutate(p_death = predict(lr_model, df_pre_laps_merging, type = "response")) |> 
  mutate(
    high_risk = case_when(
      p_death >= 0.06 ~ 1,
      TRUE ~ 0
    )) %>%
  dplyr::select(hospitalization_id, p_death, high_risk)

```

# LAPS2 Score Part 4: Prep laps2 labs ALL
```{r}
##~~~~~~~~~~~~~~~~~~~~~
# LAPS2 time
##~~~~~~~~~~~~~~~~~~~~~
 
df_encounter_laps2_temp2 <- df_encounter_laps2_temp1[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]
labs_time_dependent_2 <- labs_time_dependent[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]

df_laps2_prep1 <- 
  df_encounter_laps2_temp2 |> 
  dplyr::select(hospitalization_id, admission_dttm, discharge_dttm) |> 
  distinct() |> 
  # slice_head(n=1000) |> 
  left_join(labs_time_dependent_2 |> 
              filter(lab_category %in% c("albumin",
                                     "anion_gap",
                                     "bilirubin_total",
                                     "bun",
                                     "bicarbonate",
                                     "creatinine",
                                     "glucose_serum",
                                     # "hematocrit",
                                     "hemoglobin",
                                     "lactate",
                                     "pco2_arterial",
                                     "po2_arterial",
                                     "ph_arterial",
                                     "platelet_count",
                                     "so2_arterial",
                                     "sodium",
                                     "troponin_i",
                                     "wbc")) |> 
              dplyr::select(-lab_order_dttm),
            
            # join only if between the first 24 hours
            by = join_by(hospitalization_id)) |> 
  # making all hemoglobin into hematocrits
  mutate(lab_value_numeric = fifelse(lab_category == "hemoglobin", lab_value_numeric*3, lab_value_numeric)) |> 
  mutate(lab_category = fifelse(lab_category == "hemoglobin", "hematocrit", lab_category)) |> 
  mutate(lab_category = fifelse(lab_category == "lactate", "lactic_acid", lab_category)) |> 

    # using obs lets you get away with duplicates when going wider and can fix later!
  group_by(hospitalization_id, recorded_dttm, lab_category) |> 
  mutate(obs = row_number()) |> 
  ungroup() |> 
  
  # pivot wider to get columns
  pivot_wider(
    id_cols = c(hospitalization_id, recorded_dttm, obs),
    names_from = lab_category,
    values_from = lab_value_numeric,
    # names_prefix = "lab_wide_",
    # values_fill = NA
  ) |>
  
  # Quick min/max for duplicates at the SAME time
  dplyr::select(-obs) |> 
  mutate(lab_date = date(recorded_dttm)) |> 
  distinct() 

#ni_tic()
# find the duplicates by day
df_laps2_prep2 <- df_laps2_prep1 |>
  # arrange(clif_hospitalizations_joined_id, recorded_dttm) |>
  group_by(hospitalization_id, lab_date) |>
  # mutate(n = n()) |>
  # filter(n > 1) |>
  summarise(
    albumin             = fmin(albumin, na.rm = TRUE),
    anion_gap           = fmax(anion_gap, na.rm = TRUE),
    bilirubin_total     = fmax(bilirubin_total, na.rm = TRUE),
    bun                 = fmax(bun, na.rm = TRUE),
    bicarbonate         = fmin(bicarbonate, na.rm = TRUE),
    creatinine          = fmax(creatinine, na.rm = TRUE),
    glucose_serum       = fmin(glucose_serum, na.rm = TRUE),
    hematocrit          = fmax(hematocrit, na.rm = TRUE),
    lactic_acid         = fmax(lactic_acid, na.rm = TRUE),
    pco2_arterial       = fmax(pco2_arterial, na.rm = TRUE),
    po2_arterial        = fmax(po2_arterial, na.rm = TRUE),
    ph_arterial         = fmin(ph_arterial, na.rm = TRUE),
    platelet_count      = fmin(platelet_count, na.rm = TRUE),
    so2_arterial        = fmin(so2_arterial, na.rm = TRUE),
    sodium              = fmin(sodium, na.rm = TRUE),
    troponin            = fmax(troponin_i, na.rm = TRUE),
    wbc                 = fmin(wbc, na.rm = TRUE
    ), 

    .groups = "drop"
  ) |> 
  
    # bun_cr_ratio
  mutate(
    bun_cr_ratio = bun / creatinine
  ) |> 
  
    mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 
  # get rid of duplicate rows
  distinct() 

```

# LAPS2 Score Part 5: Labs day one for cohort
```{r}
df_laps2_dayone_prep1 <- 
  df_encounter_laps2_temp2 |> 
  dplyr::select(hospitalization_id, admission_dttm, discharge_dttm, dt_24hours_after_admit) |> 
  distinct() |> 
  left_join(labs_time_dependent_2 |> 
              filter(lab_category %in% c("albumin",
                                     "anion_gap",
                                     "bilirubin_total",
                                     "bun",
                                     "bicarbonate",
                                     "creatinine",
                                     "glucose_serum",
                                     # "hematocrit",
                                     "hemoglobin",
                                     "lactate",
                                     "pco2_arterial",
                                     "po2_arterial",
                                     "ph_arterial",
                                     "platelet_count",
                                     "so2_arterial",
                                     "sodium",
                                     "troponin_i",
                                     "wbc")) |>            
              dplyr::select(-lab_order_dttm),
            
            # join only if between the first 24 hours
            by = join_by(hospitalization_id, 
                         dt_24hours_after_admit >= recorded_dttm)
  ) |> 

  # making all hemoglobin into hematocrits
  mutate(lab_value_numeric = fifelse(lab_category == "hemoglobin", lab_value_numeric*3, lab_value_numeric)) |> 
  mutate(lab_category = fifelse(lab_category == "hemoglobin", "hematocrit", lab_category)) |> 
  mutate(lab_category = fifelse(lab_category == "lactate", "lactic_acid", lab_category)) |> 
  
  group_by(hospitalization_id, recorded_dttm, lab_category) |> 
  # using obs lets you get away with duplicates when going wider and can fix later!
  mutate(obs = row_number()) |> 
  ungroup() |> 
  
  # pivot wider to get columns
  pivot_wider(
    id_cols = c(hospitalization_id, recorded_dttm, obs, admission_dttm),
    names_from = lab_category,
    values_from = lab_value_numeric,
    # names_prefix = "lab_wide_",
    # values_fill = NA
  ) |>
  
  # Quick min/max for duplicates at the SAME time
  dplyr::select(-obs) |> 
  distinct() 

# get min/max
df_laps2_dayone_prep2 <- df_laps2_dayone_prep1 |>
  arrange(hospitalization_id, recorded_dttm) |> 
  
  # filter out those without a recorded_dttm so we don't add a date later
  filter(!is.na(recorded_dttm)) |> 
  # if we make "lab_date" the encounter date... then we are switching all labs within 24hr of admission to the "day 1" in a sneaky fashion
  mutate(lab_date = date(admission_dttm)) |> # make it all the same day 1
  group_by(hospitalization_id, lab_date) |>
  # mutate(n = n()) |>
  # filter(n > 1) |>
  summarise(
    albumin             = fmin(albumin, na.rm = TRUE),
    anion_gap           = fmax(anion_gap, na.rm = TRUE),
    bilirubin_total     = fmax(bilirubin_total, na.rm = TRUE),
    bun                 = fmax(bun, na.rm = TRUE),
    bicarbonate         = fmin(bicarbonate, na.rm = TRUE),
    creatinine          = fmax(creatinine, na.rm = TRUE),
    glucose_serum       = fmin(glucose_serum, na.rm = TRUE),
    hematocrit          = fmax(hematocrit, na.rm = TRUE),
    lactic_acid         = fmax(lactic_acid, na.rm = TRUE),
    pco2_arterial       = fmax(pco2_arterial, na.rm = TRUE),
    po2_arterial        = fmax(po2_arterial, na.rm = TRUE),
    ph_arterial         = fmin(ph_arterial, na.rm = TRUE),
    platelet_count      = fmin(platelet_count, na.rm = TRUE),
    so2_arterial        = fmin(so2_arterial, na.rm = TRUE),
    sodium              = fmin(sodium, na.rm = TRUE),
    troponin            = fmax(troponin_i, na.rm = TRUE),
    wbc                 = fmin(wbc, na.rm = TRUE
    ),

    .groups = "drop"
  ) |> 
  
    # bun_cr_ratio
  mutate(
    bun_cr_ratio = bun / creatinine
  ) |> 
  
    # replace inf with NAs
  mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 

    # get rid of duplicate rows
  distinct() 



# switch out the day 1 stuff 
df_laps2_final <- df_laps2_prep2 |> anti_join(df_laps2_dayone_prep2 |> 
                              dplyr::select(hospitalization_id, lab_date)) |> 
  bind_rows(df_laps2_dayone_prep2) |> 
  arrange(hospitalization_id, lab_date) |> 
  

  # filter out NA for lab date
  filter(!is.na(lab_date))

  
```


# LAPS2 Score Part 6: GCS Fix
```{r}
df_gcs_merge_ready <- pat_assess_gcs_time_dependent_wide |>
  mutate(vital_value = case_when(
    gcs_total %in% c(14, 15)                ~ 1,
    gcs_total %in% c(8, 9, 10, 11, 12, 13)  ~ 3,
    gcs_total %in% c(2, 3, 4, 5, 6, 7)      ~ 4,
    TRUE                                    ~ NA   # we account for this when doing the points later
  )) |>
  mutate(vital_category = "gcs") |> 
  dplyr::select(hospitalization_id,
         vital_category,
         recorded_dttm,
         vital_value,
         )


df_vitals_gcs <- vitals_time_dependent |>
              filter(vital_category %in% c("temp_c",
                                       "sbp",
                                       "spo2",
                                       "heart_rate",
                                       "respiratory_rate")) |>
              # dplyr::select(-meas_site_name) |> 
              
                          # get gcs in there now
              bind_rows(df_gcs_merge_ready)


```

# LAPS2 Score Part 7: Combine Vitals All and GCS
```{r}

df_laps2_vitals_prep1 <- 
  df_encounter_laps2_temp1 |> 
  dplyr::select(hospitalization_id, admission_dttm, discharge_dttm) |>
  distinct() |> 
  left_join(df_vitals_gcs,
            
            # join by the encounter ID
            by = join_by(hospitalization_id)) |> 
  
    group_by(hospitalization_id, recorded_dttm, vital_category) |> 
  # using obs lets you get away with duplicates when going wider and can fix later!
  mutate(obs = row_number()) |> 
  ungroup() |> 
  pivot_wider(
    id_cols = c(hospitalization_id, recorded_dttm, obs),
    names_from = vital_category,
    values_from = vital_value,
    # names_prefix = "vital_wide_",
    # values_fill = NA
  ) |>
  
  #fixing temp 
  mutate(temp = (temp_c * (9/5)) + 32) |> 
  
  #renaming resp
  rename(resp = respiratory_rate) |> 
  
  # Quick min/max for duplicates at the SAME time
  dplyr::select(-obs) |> 
  distinct() 

# find the duplicates by day
df_laps2_vitals_prep2 <- df_laps2_vitals_prep1 |>
  arrange(hospitalization_id, recorded_dttm) |>
  mutate(vital_date = date(recorded_dttm)) |> 
  group_by(hospitalization_id, vital_date) |>
  # mutate(n = n()) |>
  # filter(n > 1) |>
  summarise(
    temp                = fmin(temp, na.rm = TRUE),
    sbp                 = fmin(sbp, na.rm = TRUE),
    spo2                = fmin(spo2, na.rm = TRUE),
    heart_rate               = fmax(heart_rate, na.rm = TRUE),
    resp                = fmax(resp, na.rm = TRUE),
    gcs                 = fmax(gcs, na.rm = TRUE),  # remember the groupings... higher is lower gcs
    # shock_index         = fmax(shock_index, na.rm = TRUE),  # do this after

    .groups = "drop"
  ) |> 
  mutate(
    shock_index = heart_rate / sbp
  ) |> 
  # replace inf with NAs
  mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 

  # get rid of duplicate rows
  distinct() 

```

# LAPS2 Score Part 8: Day One Vitals for Cohort
```{r}
df_laps2_vitals_dayone_prep1 <- 
  df_encounter_laps2_temp1 |> 
  dplyr::select(hospitalization_id, admission_dttm, discharge_dttm, dt_24hours_after_admit) |>
  distinct() |> 
  left_join(df_vitals_gcs,
            
            # join only if between the first 24 hours
            by = join_by(hospitalization_id, 
                         dt_24hours_after_admit >= recorded_dttm)
  ) |> 
  
    group_by(hospitalization_id, recorded_dttm, vital_category) |> 
  # using obs lets you get away with duplicates when going wider and can fix later!
  mutate(obs = row_number()) |> 
  ungroup() |> 
  pivot_wider(
    id_cols = c(hospitalization_id, recorded_dttm, obs, admission_dttm),
    names_from = vital_category,
    values_from = vital_value,
    # names_prefix = "vital_wide_",
    # values_fill = NA
  ) |>
  
  #fixing temp 
  mutate(temp = (temp_c * (9/5)) + 32) |> 
  
  #renaming resp
  rename(resp = respiratory_rate) |> 

  # Quick min/max for duplicates at the SAME time
  dplyr::select(-obs) |> 

  distinct() 

# find the duplicates by day
df_laps2_vitals_dayone_prep2 <- df_laps2_vitals_dayone_prep1 |>
  arrange(hospitalization_id, recorded_dttm) |>
  
  # filter out those without a recorded_dttm so we don't add a date later
  filter(!is.na(recorded_dttm)) |> 
  
  # if we make "lab_date" the encounter date... then we are switching all labs within 24hr of admission to the "day 1" in a sneaky fashion
  #using encounter admit date to collect everything together
  mutate(vital_date = date(admission_dttm)) |> 
  group_by(hospitalization_id, vital_date) |>
  # mutate(n = n()) |>
  # filter(n > 1) |>
  summarise(
    temp                = fmin(temp, na.rm = TRUE),
    sbp                 = fmin(sbp, na.rm = TRUE),
    spo2                = fmin(spo2, na.rm = TRUE),
    heart_rate          = fmax(heart_rate, na.rm = TRUE),
    resp                = fmax(resp, na.rm = TRUE),
    gcs                 = fmax(gcs, na.rm = TRUE),
    # shock_index         = fmax(shock_index, na.rm = TRUE),  # do this after

    .groups = "drop"
  ) |> 
  mutate(
    shock_index         = heart_rate / sbp
  ) |> 
  
  # replace inf with NAs
  mutate(across(where(is.numeric) & !c(hospitalization_id), ~ na_if(.x, is.infinite(.x)))) |> 

  # get rid of duplicate rows
  distinct() 

# switch out the day 1 stuff 
df_laps2_vitals_final <- df_laps2_vitals_prep2 |> anti_join(df_laps2_vitals_dayone_prep2 |> 
                              dplyr::select(hospitalization_id, vital_date)) |> 
  bind_rows(df_laps2_vitals_dayone_prep2) |> 
  arrange(hospitalization_id, vital_date) |> 
  
  # drop rows with no recorded date
  filter(!is.na(vital_date))

```

# LAPS 2 Score Part 9: Calculate Laps2 Score
```{r}
setDT(df_pre_laps_final)
df_pre_laps_final_cohort <- df_pre_laps_final[hospitalization_id %in% part_eleven_imv_runs$hospitalization_id]

df_laps_calc <- df_laps2_final |> 
  # get same naming for date
  rename(recorded_date = lab_date) |> 
  full_join(df_laps2_vitals_final |> 
              rename(recorded_date = vital_date),
            join_by(hospitalization_id, recorded_date)
            ) |> 
  full_join(df_pre_laps_final_cohort)
  
clif_laps2_scores <- df_laps_calc |> 
  group_by(hospitalization_id, recorded_date) |> 
  
  mutate(laps2 = 0) |> 
  
  mutate(
    # lactic_acid and pH 
    laps2 = fcase(
      
      # missing data
      is.na(ph_arterial)  & is.na(lactic_acid) & high_risk == 0 , laps2 + 0,
      is.na(ph_arterial)  & is.na(lactic_acid) & high_risk == 1 , laps2 + 15,
      is.na(ph_arterial)  & lactic_acid < 2    & high_risk == 0 , laps2 + 0,
      is.na(ph_arterial)  & lactic_acid < 2    & high_risk == 1 , laps2 + 12,
      is.na(ph_arterial)  & lactic_acid >= 2   & lactic_acid < 4    & high_risk == 0 , laps2 + 12,
      is.na(ph_arterial)  & lactic_acid >= 2   & lactic_acid < 4    & high_risk == 1 , laps2 + 15,
      is.na(ph_arterial)  & lactic_acid >= 4   & high_risk == 0 , laps2 + 26,
      is.na(ph_arterial)  & lactic_acid >= 4   & high_risk == 1 , laps2 + 30,
      ph_arterial < 7.2   & is.na(lactic_acid) & high_risk == 0 , laps2 + 13,
      ph_arterial < 7.2   & is.na(lactic_acid) & high_risk == 1 , laps2 + 19,
      ph_arterial >= 7.2  & ph_arterial < 7.35 & is.na(lactic_acid) & high_risk == 0 , laps2 + 5,
      ph_arterial >= 7.2  & ph_arterial < 7.35 & is.na(lactic_acid) & high_risk == 1 , laps2 + 15,
      ph_arterial >= 7.35 & ph_arterial < 7.45 & is.na(lactic_acid) & high_risk == 0 , laps2 + 0,
      ph_arterial >= 7.35 & ph_arterial < 7.45 & is.na(lactic_acid) & high_risk == 1 , laps2 + 12,
      ph_arterial >= 7.45 & is.na(lactic_acid) & high_risk == 0 , laps2 + 12,
      ph_arterial >= 7.45 & is.na(lactic_acid) & high_risk == 1 , laps2 + 15,

      # complete data
      ph_arterial < 7.2 &   lactic_acid <  2, laps2 + 13,
      ph_arterial < 7.2 &   lactic_acid >= 2 & lactic_acid <  4, laps2 + 19,
      ph_arterial < 7.2 &   lactic_acid >= 4, laps2 + 34,
      ph_arterial >= 7.2 &  ph_arterial < 7.35 & lactic_acid <  2, laps2 + 5,
      ph_arterial >= 7.2 &  ph_arterial < 7.35 & lactic_acid >= 2 & lactic_acid < 4, laps2 + 15,
      ph_arterial >= 7.2 &  ph_arterial < 7.35 & lactic_acid >= 4, laps2 + 25,
      ph_arterial >= 7.35 & ph_arterial < 7.45 & lactic_acid <  2, laps2 + 0,
      ph_arterial >= 7.35 & ph_arterial < 7.45 & lactic_acid >= 2 & lactic_acid < 4, laps2 + 12,
      ph_arterial >= 7.35 & ph_arterial < 7.45 & lactic_acid >= 4, laps2 + 26,
      ph_arterial >= 7.45 & lactic_acid <  2, laps2 + 12,
      ph_arterial >= 7.45 & lactic_acid >= 2   & lactic_acid <  4, laps2 + 15,
      ph_arterial >= 7.45 & lactic_acid >= 4, laps2 + 30,
      default                                                                       = laps2
      ),
    
    # Sodium
    laps2 = fcase(
      is.na(sodium), laps2 + 0,
      sodium <  129, laps2 + 14,
      sodium >= 129 & sodium < 135, laps2 + 7,
      sodium >= 135 & sodium < 146, laps2 + 0,
      sodium >= 146, laps2 + 4,
      default = laps2
    ),
    
    # Bilirubin
    laps2 = fcase(
      is.na(bilirubin_total), laps2 + 0,
      bilirubin_total <  2, laps2 + 0,
      bilirubin_total >= 2 & bilirubin_total < 3, laps2 + 11,
      bilirubin_total >= 3 & bilirubin_total < 5, laps2 + 18,
      bilirubin_total >= 5 & bilirubin_total < 8, laps2 + 25,
      bilirubin_total >= 8, laps2 + 41,
      TRUE , laps2
    ),
    
    # BUN
    laps2 = fcase(
      is.na(bun), laps2 + 0,
      bun < 18, laps2 + 0,
      bun >= 18 & bun < 20, laps2 + 11,
      bun >= 20 & bun < 40, laps2 + 12,
      bun >= 40 & bun < 80, laps2 + 20,
      bun >= 80, laps2 + 25,
      default = laps2
    ),
    
    # Creatinine
    laps2 = fcase(
      is.na(creatinine), laps2 + 0,
      creatinine < 1, laps2 + 0,
      creatinine >= 1 & creatinine < 2, laps2 + 6,
      creatinine >= 2 & creatinine < 4, laps2 + 11,
      creatinine >= 4, laps2 + 5,
      default = laps2
    ),
    
    # BUN/Cr Ratio
    laps2 = fcase(
      is.na(bun_cr_ratio), laps2 + 0,
      bun_cr_ratio < 25, laps2 + 0,
      bun_cr_ratio >= 25, laps2 + 10,
      default = laps2
    ),
    
    # Albumin
    laps2 = fcase(
      is.na(albumin), laps2 + 0,
      albumin < 2, laps2 + 31,
      albumin >= 2 & albumin < 2.5, laps2 + 15,
      albumin >= 2.5, laps2 + 0,
      default = laps2
    ),
    
    # glucose_serum
    laps2 = fcase(
      is.na(glucose_serum), laps2 + 0,
      glucose_serum < 40, laps2 + 10,
      glucose_serum >= 40 & glucose_serum < 60, laps2 + 10,
      glucose_serum >= 60 & glucose_serum < 200, laps2 + 0,
      glucose_serum >= 200, laps2 + 3,
      default                                                                   = laps2
    ),
    
    # Hematocrit (Hct)
    laps2 = fcase(
      is.na(hematocrit), laps2 + 0,
      hematocrit < 20, laps2 + 7,
      hematocrit >= 20 & hematocrit < 40, laps2 + 8,
      hematocrit >= 40 & hematocrit < 50, laps2 + 0,
      hematocrit >= 50, laps2 + 3,
      default = laps2
    ),
    
    # WBC
    laps2 = fcase(
      is.na(wbc) & high_risk == 0 , laps2 + 0,
      is.na(wbc) & high_risk == 1 , laps2 + 32,
      wbc < 5, laps2 + 8,
      wbc >= 5 & wbc < 13, laps2 + 0,
      wbc >= 13, laps2 + 11,
      default = laps2
    ),
    
  # pco2_arterial
    laps2 = fcase(
      is.na(pco2_arterial), laps2 + 0,
      pco2_arterial < 35, laps2 + 7,
      pco2_arterial >= 35 & pco2_arterial < 45, laps2 + 0,
      pco2_arterial >= 45 & pco2_arterial < 55, laps2 + 11,
      pco2_arterial >= 55 & pco2_arterial < 65, laps2 + 13,
      pco2_arterial >= 65, laps2 + 12,
      default = laps2
    ),
  
  # po2_arterial
    laps2 = fcase(
      is.na(po2_arterial), laps2 + 0,
      po2_arterial < 50, laps2 + 8,
      po2_arterial >= 50 & po2_arterial < 120, laps2 + 0,
      po2_arterial >= 120, laps2 + 12,
      default = laps2
    ),
  
  # Troponin
    laps2 = fcase(
      is.na(troponin)  & high_risk == 0 , laps2 + 0,
      is.na(troponin)  & high_risk == 1 , laps2 + 9,
      troponin <  0.01, laps2 + 0,
      troponin >= 0.01 & troponin < 0.2, laps2 + 8,
      troponin >= 0.2  & troponin < 1, laps2 + 17,
      troponin >= 1    & troponin < 3, laps2 + 19,
      troponin >= 3, laps2 + 25,
      default = laps2
    ),
  
  # Temp
    laps2 = fcase(
      is.na(temp), laps2 + 0,
      temp < 96, laps2 + 20,
      temp >= 96    & temp < 100.5, laps2 + 0,
      temp >= 100.5, laps2 + 3,
      default = laps2
    ),

  # heart_rate
    laps2 = fcase(
      is.na(heart_rate), laps2 + 0,
      heart_rate < 60, laps2 + 7,
      heart_rate >= 60  & heart_rate < 110, laps2 + 0,
      heart_rate >= 110 & heart_rate < 140, laps2 + 7,
      heart_rate >= 140, laps2 + 10,
      default = laps2
    ),
  
  # Resp
    laps2 = fcase(
      is.na(resp), laps2 + 0,
      resp < 20, laps2 + 0,
      resp >= 20 & resp < 30, laps2 + 11,
      resp >= 30, laps2 + 21,
      default = laps2
    ),
    
  # SBP
    laps2 = fcase(
      is.na(sbp), laps2 + 0,
      sbp < 75, laps2 + 22,
      sbp >= 75  & sbp < 90, laps2 + 13,
      sbp >= 90  & sbp < 120, laps2 + 5,
      sbp >= 120 & sbp < 140, laps2 + 0,
      sbp >= 140 & sbp < 160, laps2 + 8,
      sbp >= 160, laps2 + 14,
      default = laps2
    ),
  
  # Shock
    laps2 = fcase(
      is.na(shock_index), laps2 + 0,
      shock_index < 0.65, laps2 + 0,
      shock_index >= 0.65 & shock_index < 0.85, laps2 + 8,
      shock_index >= 0.85, laps2 + 17,
      default = laps2
    ),
  
  # O2Sat
    laps2 = fcase(
      is.na(spo2) & high_risk == 0 , laps2 + 0,
      is.na(spo2) & high_risk == 1 , laps2 + 22,
      spo2 <  90, laps2 + 22,
      spo2 >= 90  & spo2 < 94, laps2 + 12,
      spo2 >  94, laps2 + 0,
      default = laps2
    ),
  
  # Neuro (GCS)
    laps2 = fcase(
      is.na(gcs) & high_risk == 0 , laps2 + 16,
      is.na(gcs) & high_risk == 1 , laps2 + 21,
      gcs == 1, laps2 + 0,
      gcs == 2, laps2 + 16,
      gcs == 3, laps2 + 21,
      gcs == 4, laps2 + 36,
      default = laps2
    )
  )

```

# LAPS2 Score Final:
```{r}

pre_final_cohort_imv_runs <-  part_eleven_imv_runs |>
  mutate(imv_date = date(begin_imv))

begin_imv_laps2 <- merge(pre_final_cohort_imv_runs, clif_laps2_scores, by = "hospitalization_id", all = TRUE)

final_cohort_imv_runs <- begin_imv_laps2 %>%
  filter(imv_date == recorded_date)

#LAPS2 MISSING DATA SUMMARY 

# Step 1: Flag hospitalizations missing data for each SOFA component
laps2_missing_summary <- final_cohort_imv_runs %>%
  group_by(hospitalization_id) %>%
  summarise(
    missing_sodium = all(is.na(sodium)),
    missing_bili = all(is.na(bilirubin_total)),
    missing_bun = all(is.na(bun)),
    missing_creat = all(is.na(creatinine)),
    missing_bun_cr_ratio = all(is.na(bun_cr_ratio)),
    missing_albumin = all(is.na(albumin)),
    missing_gluc = all(is.na(glucose_serum)),
    missing_hct = all(is.na(hematocrit)),
    missing_wbc = all(is.na(wbc)),
    missing_pco2 = all(is.na(pco2_arterial)),
    missing_po2 = all(is.na(po2_arterial)),
    missing_trop = all(is.na(troponin)),
    missing_temp = all(is.na(temp)),
    missing_hr = all(is.na(heart_rate)),
    missing_resp = all(is.na(resp)),
    missing_sbp = all(is.na(sbp)),
    missing_shock_index = all(is.na(shock_index)),
    missing_spo2 = all(is.na(spo2)),
    missing_gcs = all(is.na(gcs)),
    .groups = "drop"
  )

# Step 2: Calculate percent missing
laps2_missing_percent <- laps2_missing_summary %>%
  summarise(
    pct_missing_sodium = mean(missing_sodium) * 100,
    pct_missing_bili = mean(missing_bili) * 100,
    pct_missing_bun = mean(missing_bun) * 100,
    pct_missing_creat = mean(missing_creat) * 100,
    pct_missing_bun_cr_ratio = mean(missing_bun_cr_ratio) * 100,
    pct_missing_albumin = mean(missing_albumin) * 100,
    pct_missing_gluc = mean(missing_gluc) * 100,
    pct_missing_hct = mean(missing_hct) * 100,
    pct_missing_wbc = mean(missing_wbc) * 100,
    pct_missing_pco2 = mean(missing_pco2) * 100,
    pct_missing_po2 = mean(missing_po2) * 100,
    pct_missing_trop = mean(missing_trop) * 100,
    pct_missing_temp = mean(missing_temp) * 100,
    pct_missing_hr = mean(missing_hr) * 100,
    pct_missing_resp = mean(missing_resp) * 100,
    pct_missing_sbp = mean(missing_sbp) * 100,
    pct_missing_shock_index = mean(missing_shock_index) * 100,
    pct_missing_spo2 = mean(missing_spo2) * 100,
    pct_missing_gcs = mean(missing_gcs) * 100
  )

final_cohort_imv_runs_reduced <- final_cohort_imv_runs[, .(hospitalization_id, begin_imv, end_imv, admission_dttm, discharge_dttm, hospital_id, icu_type, hospital_icu_id, location_of_intubation, first_location_in_hospital, first_trach_dttm, imv_duration_hours_final, imv_duration_days, age_at_admission, age_category, death_dttm, admission_type_category, discharge_category, sex_category, race_category, ethnicity_category, ethnicity_category_final, SVM_SCORE, language_category, top_five_non_eng_non_spanish, non_eng, elixhauser_score, weight_kg, bmi, trached, reintubation, thirty_day_mortality, vfd_clean, hfd_6m, laps2, deep_sedation_weighted, deep_sedation_unweighted, rass_weighted_avg, rass_below_minus_3_count, median_rass_binary, median_rass_cont, sedation_index_score, mean_gcs, median_gcs, death_during_admission, rass_frequency_hrs, icufd_28, time_to_icu_admission, hfd_28d)]

# Save the Missing results to a file
fwrite(laps2_missing_summary, file = paste0(final_data, institution, "_laps2_missing_summary.csv"))
fwrite(laps2_missing_percent, file = paste0(final_data, institution, "_laps2_missing_percentages.csv"))

#Save Point
write_parquet(final_cohort_imv_runs_reduced, paste0(intermediate_dir, "cleaned_cohort_w_laps", institution, ".parquet"))
```

# PART THIRTEEN: Add in COVID Data
```{r}
#Save Point Code if need to re-run and don't want to re-run LAPS2 can simply run parts 0- 12 to upload tables/directories then remove # before code below and start here.
#final_cohort_imv_runs_reduced <- read_parquet(paste0(intermediate_dir, "cleaned_cohort_w_laps", institution, ".parquet"))

hospitalization_covid_filt <- hospitalization[hospitalization_id %in% final_cohort_imv_runs_reduced$hospitalization_id]
setnames(hospitalization_covid_filt, "admission_dttm", "cohort_admission_dttm")
hospitalization_covid_filt[, cohort_hospitalization_id := hospitalization_id]
#setnames(hospitalization_covid_filt, "hospitalization_id", "cohort_hospitalization_id")
hospitalization_covid_filt <- hospitalization_covid_filt[, .(patient_id, hospitalization_id, cohort_hospitalization_id, cohort_admission_dttm)]

hospitalization_covid_all <- hospitalization[patient_id %in% hospitalization_covid_filt$patient_id]

hospitalization_covid_merge <- merge(hospitalization_covid_all, hospitalization_covid_filt, by = c("patient_id", "hospitalization_id"), all.x = TRUE)

hospitalization_covid_merge[, cohort_admission_dttm := nafill(cohort_admission_dttm, type = "locf"), by = patient_id]
hospitalization_covid_merge[, cohort_admission_dttm := nafill(cohort_admission_dttm, type = "nocb"), by = patient_id]
hospitalization_covid_merge[, cohort_hospitalization_id := na.locf(cohort_hospitalization_id, na.rm = FALSE), by = patient_id]
hospitalization_covid_merge[, cohort_hospitalization_id := na.locf(cohort_hospitalization_id, fromLast = TRUE, na.rm = FALSE), by = patient_id]

hospitalization_covid_merge_filt <- hospitalization_covid_merge[
  admission_dttm >= (cohort_admission_dttm %m-% months(3)) &
  admission_dttm <= cohort_admission_dttm
]
hospitalization_covid_merge_filt <- hospitalization_covid_merge_filt[, .(patient_id, hospitalization_id, cohort_hospitalization_id, admission_dttm, cohort_admission_dttm)]

covid_testing <- as.data.table(read_data(paste0(data_dir, "clif_microbiology_nonculture", file_type), select = c("hospitalization_id", "result_dttm", "result_category")))
id_cw <- hospitalization %>% distinct(hospitalization_id, patient_id)
covid_testing_pat <- covid_testing %>% left_join(id_cw, by = "hospitalization_id")

covid_testing_pat <- unique(covid_testing_pat, by = c("hospitalization_id", "result_dttm"))
setDT(covid_testing_pat)

covid_positive <- covid_testing_pat[result_category == "Detected", .(patient_id, hospitalization_id, result_dttm, result_category)]
setnames(covid_positive, "result_category", "covid_status")
covid_positive <- unique(covid_positive, by = c("hospitalization_id", "covid_status"))

covid_positive_hosp_merge <- merge(hospitalization_covid_merge_filt, covid_positive, by = c("hospitalization_id", "patient_id"), all = TRUE)
setorder(covid_positive_hosp_merge, patient_id, hospitalization_id, admission_dttm)

#LOCF
covid_positive_hosp_merge_locf <- covid_positive_hosp_merge[, `:=`(
  result_dttm = na.locf(result_dttm, na.rm = FALSE),
  covid_status = na.locf(covid_status, na.rm = FALSE)
), by = patient_id]


#Retain hospitalizations 1 month prior to cohort admission
covid_positive_hosp_merge_locf_filt <- covid_positive_hosp_merge_locf[
  admission_dttm >= (cohort_admission_dttm %m-% months(1)) &
  admission_dttm <= cohort_admission_dttm                     
]

#Filter Only for COVID Positive Near Admission
covid_positive_hosp_merge_locf_filt <- covid_positive_hosp_merge_locf_filt[
  result_dttm >= (cohort_admission_dttm %m-% months(1)) &
  result_dttm <= (cohort_admission_dttm %m+% months(1))
]

covid_positive_hosp_merge_locf_filt_final <- covid_positive_hosp_merge_locf_filt[, .(cohort_hospitalization_id, covid_status)]
setnames(covid_positive_hosp_merge_locf_filt_final, "cohort_hospitalization_id", "hospitalization_id")
covid_positive_hosp_merge_locf_filt_final <- unique(covid_positive_hosp_merge_locf_filt_final, by = c("hospitalization_id", "covid_status"))

final_cohort_imv_runs_reduced <- merge(final_cohort_imv_runs_reduced, covid_positive_hosp_merge_locf_filt_final, by = "hospitalization_id", all = TRUE)
final_cohort_imv_runs_reduced[is.na(covid_status), covid_status := "Not Detected"]
```

# PART FOURTEEN: Perform Data Transformations - Long to Wide
```{r}
# Vitals
#Filtered Vitals
time_dependent_vitals_2 <- c("sbp", "dbp", "map", "spo2")
vitals_time_dependent_2 <- vitals_time_dependent[vital_category %in% time_dependent_vitals_2 & hospitalization_id %in% final_cohort_imv_runs_reduced$hospitalization_id]

vitals_time_dependent_wide <- dcast(vitals_time_dependent_2, 
                   hospitalization_id + recorded_dttm ~ vital_category, 
                   value.var = "vital_value", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
setorder(vitals_time_dependent_wide, hospitalization_id, recorded_dttm)

# Calculate MAP values
vitals_time_dependent_wide <- vitals_time_dependent_wide %>%
  mutate(map = case_when(
    is.na(map) | map == "" | map == 0 ~ (1/3 * sbp) + (2/3 * dbp),
    TRUE ~ map
  ))
vitals_time_dependent_wide <- vitals_time_dependent_wide[, .(hospitalization_id, recorded_dttm, map, spo2)]

vitals_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]

# Labs
labs_time_dependent_wide <- dcast(labs_time_dependent_2, 
                   hospitalization_id + recorded_dttm ~ lab_category, 
                   value.var = "lab_value_numeric", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
setorder(labs_time_dependent_wide, hospitalization_id, recorded_dttm)

labs_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]

# Meds
meds_time_dependent_wide <- dcast(meds_time_dependent, 
                   hospitalization_id + recorded_dttm ~ med_category, 
                   value.var = "med_dose", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
setorder(meds_time_dependent_wide, hospitalization_id, recorded_dttm)
meds_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]

# Patient Assessments
#RASS
pat_assess_rass_time_dependent_wide <- pat_assess_rass_time_dependent
setorder(pat_assess_rass_time_dependent_wide, hospitalization_id, recorded_dttm)

#GCS
pat_assess_gcs_time_dependent_wide <- dcast(pat_assess_gcs_time_dependent, 
                   hospitalization_id + recorded_dttm ~ assessment_category, 
                   value.var = "numerical_value", 
                   fun.aggregate = mean,  # Aggregate duplicates using mean
                   fill = NA)  # Fill missing values with NA
setorder(pat_assess_gcs_time_dependent_wide, hospitalization_id, recorded_dttm)

pat_assess_gcs_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]

# Respiratory
respiratory_support_time_dependent_wide <- respiratory_support |>
  dplyr::filter(hospitalization_id %in% final_cohort_imv_runs_reduced$hospitalization_id) |>
  dplyr::select(hospitalization_id, recorded_dttm, device_category, mode_category, fio2_set, peep_set, lpm_set, tidal_volume_set) |>
  dplyr::mutate(
    fio2_set = as.numeric(fio2_set),
    lpm_set = as.numeric(lpm_set),
    fio2_set = if (mean(fio2_set, na.rm = TRUE) > 1) fio2_set / 100 else fio2_set,
    fio2_set = ifelse(fio2_set >= 0.21 & fio2_set <= 1, fio2_set, NA_real_),
    lpm_set = ifelse(lpm_set >= 0 & lpm_set <= 60, lpm_set, NA_real_)
  )

respiratory_support_time_dependent_wide[, recorded_dttm := as.POSIXct(recorded_dttm, format = "%Y-%m-%d %H:%M:%S", tz = time_zone)]

respiratory_support_time_dependent_wide <- respiratory_support_time_dependent_wide %>%
  # Fill in device_category based on other values
  dplyr::mutate(device_category = case_when(
    mode_category %in% c("SIMV", "Pressure-regulated Volume Control", "Assist Control-Volume Control") ~ "IMV",
    is.na(device_category) & fio2_set == 0.21 & is.na(lpm_set) & is.na(peep_set) & is.na(tidal_volume_set) ~ "Room Air",
    is.na(device_category) & is.na(fio2_set) & lpm_set == 0 & is.na(peep_set) & is.na(tidal_volume_set) ~ "Room Air",
    is.na(device_category) & is.na(fio2_set) & lpm_set >= 20 & is.na(peep_set) & is.na(tidal_volume_set) ~ "High Flow NC",
    device_category == "Nasal Cannula" & is.na(fio2_set) & lpm_set >= 20 ~ "High Flow NC",
    is.na(device_category) & is.na(fio2_set) & lpm_set < 20 & lpm_set > 0 & is.na(peep_set) & is.na(tidal_volume_set) ~ "Nasal Cannula",
    TRUE ~ device_category
  )) %>%
  # Create fio2_combined based on existing values
  dplyr::mutate(fio2_combined = case_when(
    !is.na(fio2_set) ~ fio2_set,
    is.na(fio2_set) & device_category == "Room Air" ~ 0.21,
    is.na(fio2_set) & device_category == "Nasal Cannula" ~ (0.24 + (0.04 * lpm_set)),
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set == 6 ~ 0.35,
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set == 7 ~ 0.41,
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set == 8 ~ 0.47,
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set == 9 ~ 0.53,
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set == 10 ~ 0.60,
    is.na(fio2_set) & device_category == "Face Mask" & lpm_set > 10 ~ 0.90,
    TRUE ~ NA_real_
  )) %>%
  # Keep only the selected columns
  dplyr::select(hospitalization_id, recorded_dttm, device_category, mode_category, lpm_set, fio2_combined)

setorder(respiratory_support_time_dependent_wide, hospitalization_id, recorded_dttm)

# List of tables to merge
tables_merge_wide <- list(vitals_time_dependent_wide,
               pat_assess_rass_time_dependent_wide,
               labs_time_dependent_wide,
               meds_time_dependent_wide, 
               pat_assess_gcs_time_dependent_wide,
               respiratory_support_time_dependent_wide)

# Iterative Merge on Hospitalization_id and recorded_dttm to form time_dependent_timeline
time_dependent_timeline_all <- reduce(tables_merge_wide, function(x, y) merge(x, y, by = c("hospitalization_id", "recorded_dttm"), all = TRUE))

```

# PART FIFTEEN: Baseline Values from 24 Hours Prior to Begin IMV
```{r}
# Begin IMV
final_cohort_timestamps <- final_cohort_imv_runs_reduced[, .(hospitalization_id, admission_dttm, discharge_dttm, begin_imv, end_imv, imv_duration_hours_final)]

# Calculate start_time (24 hours before begin_imv)
final_cohort_timestamps[, start_time := as.POSIXct(begin_imv) - hours(48), by = hospitalization_id]

# Calculate end_time based on the condition for each hospitalization_id
final_cohort_timestamps[, end_time := as.POSIXct(begin_imv), by = hospitalization_id]

#24 Hours Prior to IMV Timeline
time_dependent_timeline_pre_imv <- merge(final_cohort_timestamps, time_dependent_timeline_all, by = "hospitalization_id", all = TRUE)
setorder(time_dependent_timeline_all, hospitalization_id, recorded_dttm)

# Filter the table to retain rows where recorded_dttm is within the time range
time_dependent_timeline_pre_imv_24h <- time_dependent_timeline_pre_imv[recorded_dttm >= start_time & recorded_dttm <= end_time]
time_dependent_timeline_pre_imv_24h <- time_dependent_timeline_pre_imv_24h[, .(hospitalization_id, recorded_dttm, begin_imv, map, spo2, rass_value, bilirubin_total, creatinine, platelet_count, po2_arterial, pco2_arterial, ph_arterial, fio2_combined, gcs_total, dopamine, dobutamine, norepinephrine, epinephrine)]

#LOCF
cols_to_update_imv_24h <- c("map", "spo2", "rass_value", "bilirubin_total", "creatinine", "platelet_count", "po2_arterial", "pco2_arterial", "ph_arterial", "fio2_combined", "gcs_total", "dopamine", "dobutamine", "norepinephrine", "epinephrine")
time_dependent_timeline_pre_imv_24h[, (cols_to_update_imv_24h) := lapply(.SD, function(x) fifelse(x == "", NA, x)), .SDcols = cols_to_update_imv_24h]

time_dependent_timeline_pre_locf <- time_dependent_timeline_pre_imv_24h[, (cols_to_update_imv_24h) := lapply(.SD, function(x) na.locf(x, na.rm = FALSE)), 
                 by = hospitalization_id, .SDcols = cols_to_update_imv_24h]

# Define columns for min and max selection
cols_min_1 <- c("rass_value", "map", "spo2", "platelet_count", "po2_arterial", "ph_arterial", "gcs_total", "begin_imv")
cols_max_1 <- c("bilirubin_total", "creatinine", "pco2_arterial", "fio2_combined", "dopamine", "dobutamine", 
              "norepinephrine", "epinephrine", "recorded_dttm")

# Compute the min and max for each hospitalization_id and hour_imv
time_dependent_timeline_pre_final <- time_dependent_timeline_pre_locf[, 
    lapply(.SD, min), by = .(hospitalization_id), .SDcols = cols_min_1][
    time_dependent_timeline_pre_locf[, lapply(.SD, max), by = .(hospitalization_id), .SDcols = cols_max_1], 
    on = .(hospitalization_id)
]
 
time_dependent_timeline_last_primary <- time_dependent_timeline_pre_final

time_dependent_timeline_last_primary[, recorded_dttm := begin_imv - minutes(15)]
time_dependent_timeline_last_primary[, hour_imv := 0]
time_dependent_timeline_last_primary[, c("begin_imv") := NULL]


#Sensitivity

time_dependent_timeline_last_sensitivity <- time_dependent_timeline_last_primary
time_dependent_timeline_last_primary$rass_value <- -5
time_dependent_timeline_last_sensitivity$rass_value <- -1

```

# PART SIXTEEN: Form Cox Table
```{r}
timeline_meta <- final_cohort_timestamps[, .(
  start = min(begin_imv, na.rm=TRUE),
  dur = max(0, pmin(72, min(imv_duration_hours_final, na.rm=TRUE)))
), by = hospitalization_id]

imv_timeline_72h <- timeline_meta[, .(
  recorded_dttm = seq(
    from = start, 
    to = start + dhours(dur), # dhours() gives a lubridate duration
    by = "1 hour"
  )
), by = .(hospitalization_id)]

# Calculate start_time
imv_timeline_72h[, start_time := min(recorded_dttm, na.rm = TRUE), by = hospitalization_id]

# Calculate end_time as the last recorded_dttm for each hospitalization_id
imv_timeline_72h[, end_time := max(recorded_dttm, na.rm = TRUE), by = hospitalization_id]

imv_timeline_72h[, hour_imv := seq_len(.N) - 1, by = hospitalization_id]

start_end_time <- imv_timeline_72h[, .SD[1], by = hospitalization_id]
start_end_time <- start_end_time[, .(hospitalization_id, start_time, end_time)]

imv_timeline_72h[, c("start_time", "end_time") := NULL]

# MERGE IT UP
total_timeline <- merge(imv_timeline_72h, time_dependent_timeline_all, by = c("hospitalization_id", "recorded_dttm"), all = TRUE)
setcolorder(total_timeline, c("hospitalization_id", "hour_imv", "recorded_dttm", "rass_value", "map", "spo2", "bilirubin_total", "creatinine", "platelet_count", "po2_arterial", "pco2_arterial", "ph_arterial", "fio2_combined", "gcs_total", "dopamine", "dobutamine", "norepinephrine", "epinephrine", "cisatracurium", "isoproterenol", "milrinone", "phenylephrine", "rocuronium", "vasopressin", "vecuronium", "propofol", "ketamine", "dexmedetomidine", "lorazepam", "midazolam", "fentanyl", "hydromorphone", "morphine", "remifentanil"))

total_timeline_filt <- merge(total_timeline, start_end_time, by = "hospitalization_id", all = TRUE)


#LOCF
cols_to_update_imv_total <- c("hour_imv", "rass_value", "map", "spo2", "bilirubin_total", "creatinine", "platelet_count", "po2_arterial", "pco2_arterial", "ph_arterial", "fio2_combined", "gcs_total", "dopamine", "dobutamine", "norepinephrine", "epinephrine", "cisatracurium", "isoproterenol", "milrinone", "phenylephrine", "rocuronium", "vasopressin", "vecuronium", "pao2_imputed", "propofol", "ketamine", "dexmedetomidine", "lorazepam", "midazolam", "fentanyl", "hydromorphone", "morphine", "remifentanil")

total_timeline_primary <- merge(total_timeline_filt, time_dependent_timeline_last_primary, by = c("hospitalization_id", "recorded_dttm", "hour_imv", "rass_value", "map", "spo2", "platelet_count", "po2_arterial", "ph_arterial", "gcs_total", "bilirubin_total", "creatinine", "pco2_arterial", "fio2_combined", "dopamine", "dobutamine", "norepinephrine", "epinephrine"), all = TRUE)

total_timeline_primary[, pao2_imputed := {
  s <- spo2 / 100
  a <- 11700 / ((1 / s) - 1)
  b <- sqrt(50^3 + a^2)
  (b + a)^(1/3) + (b - a)^(1/3)
}]

total_timeline_primary[, (cols_to_update_imv_total) := lapply(.SD, function(x) fifelse(x == "", NA, x)), .SDcols = cols_to_update_imv_total]

# Paralytics
total_timeline_primary[, paralytics := fifelse(rocuronium > 0 | vecuronium > 0.09 | cisatracurium > 0.49, TRUE, FALSE)]
total_timeline_primary[is.na(paralytics), paralytics := FALSE]

#Fix Logically Inconsistent RASS Values During Paralytic Periods
total_timeline_primary[
  rass_value > -3 & paralytics == "TRUE",
  rass_value := NA
]

total_timeline_locf_primary <- total_timeline_primary[, (cols_to_update_imv_total) := lapply(.SD, function(x) na.locf(x, na.rm = FALSE)), 
                 by = hospitalization_id, .SDcols = cols_to_update_imv_total]

# Filter the table to retain rows where recorded_dttm is within the time range
total_timeline_locf_primary_filt <- total_timeline_locf_primary[recorded_dttm >= start_time & recorded_dttm <= end_time]

# Define columns for min and max selection
cols_min <- c("rass_value", "map", "spo2", "platelet_count", "po2_arterial", "ph_arterial", "pao2_imputed", "gcs_total")
cols_max <- c("bilirubin_total", "creatinine", "pco2_arterial", "fio2_combined", "dopamine", "dobutamine", "norepinephrine", "epinephrine", "cisatracurium", "isoproterenol", "milrinone", "phenylephrine", "rocuronium", "vasopressin", "vecuronium", "propofol", "ketamine", "dexmedetomidine", "lorazepam", "midazolam", "fentanyl", "hydromorphone", "morphine", "remifentanil")

# Compute the min and max for each hospitalization_id and hour_imv
total_timeline_hourly_primary <- total_timeline_locf_primary_filt[, 
    lapply(.SD, min), by = .(hospitalization_id, hour_imv), .SDcols = cols_min][
    total_timeline_locf_primary_filt[, lapply(.SD, max), by = .(hospitalization_id, hour_imv), .SDcols = cols_max], 
    on = .(hospitalization_id, hour_imv)
]

#Calculate Morphine Equivalents
# ---- Function ---------------------------------------------------------------
convert_opioids_to_mme <- function(dt,
                                   conv = list(                # default factors*
                                     morphine      = 2.5,        # mg IV morphine  mg MME
                                     hydromorphone = 5,     # 1mg Dilaudid = 5
                                     fentanyl      = 0.1667,      # 1mcg fent = 0.1667
                                     remifentanil  = 0.1667
                                   ),
                                   add_total = TRUE,
                                   total_name = "mme_total") {
  ## Safety: work on data.table
  if (!is.data.table(dt)) dt <- as.data.table(dt)

  ## Identify which of the target columns are actually present
  opioid_cols <- intersect(names(conv), names(dt))
  if (length(opioid_cols) == 0) {
    warning("No opioid dose columns found; nothing to convert.")
    return(dt)
  }

  ## Create one new column per opioid with *_mme suffix
  for (col in opioid_cols) {
    new_col <- paste0(col, "_mme")
    dt[, (new_col) := get(col) * conv[[col]]]
  }

  ## Rowlevel total
if (add_total) {
  mme_cols <- paste0(opioid_cols, "_mme")
  dt[, (total_name) := {
    row_vals <- .SD
    total <- rowSums(row_vals, na.rm = TRUE)
    all_na <- apply(row_vals, 1, function(x) all(is.na(x)))
    ifelse(all_na, NA_real_, total)
  }, .SDcols = mme_cols]
}

  invisible(dt)
}
# ------------------------------------------------------------------------------
convert_opioids_to_mme(total_timeline_hourly_primary)
total_timeline_hourly_primary[mme_total == 0, mme_total := NA]

#Determine Average MMEs
# Step 1: Filter to hospitalization_ids that have at least one non-NA total_mme
mme_ids <- total_timeline_hourly_primary[!is.na(mme_total), unique(hospitalization_id)]

# Step 2: Subset to just those hospitalization_ids and calculate average_mme
median_mme_by_hosp <- total_timeline_hourly_primary[
  hospitalization_id %in% mme_ids,
  .(mme_median = median(mme_total, na.rm = TRUE)),
  by = hospitalization_id
]

#Calculate Benzo Equivalents
convert_benzos_to_lorazepam <- function(dt,
                                        conv = list(
                                          lorazepam  = 1,    # reference
                                          midazolam = 0.5 # 1mg midaz=0.5mg
                                          # add diazepam = 0.1, etc., if needed
                                        ),
                                        suffix = "_loraz_equiv",
                                        add_total = TRUE,
                                        total_name = "loraz_equiv_total") {
  if (!is.data.table(dt)) dt <- as.data.table(dt)

  benzo_cols <- intersect(names(conv), names(dt))
  if (!length(benzo_cols)) {
    warning("No benzodiazepine dose columns found; nothing converted.")
    return(dt)
  }

  # Create new *_loraz_equiv columns
  for (col in benzo_cols) {
    new_col <- paste0(col, suffix)
    dt[, (new_col) := get(col) * conv[[col]]]
  }

  # Optional rowlevel total
  if (add_total) {
  equiv_cols <- paste0(benzo_cols, suffix)
  dt[, (total_name) := {
    row_vals <- .SD
    total <- rowSums(row_vals, na.rm = TRUE)
    all_na <- apply(row_vals, 1, function(x) all(is.na(x)))
    ifelse(all_na, NA_real_, total)
  }, .SDcols = equiv_cols]
}

  invisible(dt)
}
convert_benzos_to_lorazepam(total_timeline_hourly_primary)
total_timeline_hourly_primary[loraz_equiv_total == 0, loraz_equiv_total := NA]

#Determine Average Benzos
# Step 1: Filter to hospitalization_ids that have at least one non-NA loraz_equiv_total
benzo_ids <- total_timeline_hourly_primary[!is.na(loraz_equiv_total), unique(hospitalization_id)]

# Step 2: Subset to just those hospitalization_ids and calculate average_benzos
median_benzo_by_hosp <- total_timeline_hourly_primary[
  hospitalization_id %in% benzo_ids,
  .(loraz_median = median(loraz_equiv_total, na.rm = TRUE)),
  by = hospitalization_id
]

#Lets Calculate Average Propofol
# Step 1: Identify hospitalization_ids where all propofol values are 0 (or NA/0)
setDT(total_timeline_hourly_primary)
zero_only_ids <- total_timeline_hourly_primary[
  , .(all_zero = all(propofol == 0, na.rm = TRUE)),
  by = hospitalization_id
][all_zero == TRUE, hospitalization_id]

# Step 2: Set propofol := NA for those hospitalization_ids
total_timeline_hourly_primary[
  hospitalization_id %in% zero_only_ids,
  propofol := NA
]

# Step 1: Filter to hospitalization_ids that have at least one non-NA
propofol_ids <- total_timeline_hourly_primary[!is.na(propofol), unique(hospitalization_id)]

# Step 2: Subset to just those hospitalization_ids and calculate average
median_propofol_by_hosp <- total_timeline_hourly_primary[
  hospitalization_id %in% propofol_ids,
  .(propofol_median = median(propofol, na.rm = TRUE)),
  by = hospitalization_id
]

##Opioid Only Sedation
# Define sedative and opioid columns
sedative_cols <- c("propofol", "loraz_equiv_total", "dexmedetomidine", "pentobarbital", "ketamine", "phenobarbital")
mme_col <- "mme_total"

# Create sedative indicators (NA or 0)
for (col in sedative_cols) {
  if (!col %in% names(total_timeline_hourly_primary)) {
    # Column missing  create it filled with NA
    total_timeline_hourly_primary[, (col) := NA_real_]
  }
}

# Ensure mme_total column exists
if (!mme_col %in% names(total_timeline_hourly_primary)) {
  stop("mme_total column is missing  cannot proceed.")
}

# Summarize per hospitalization_id
opioid_only_summary <- total_timeline_hourly_primary[
  , .(
      all_sedatives_na_or_zero = all(
        is.na(propofol) | propofol == 0
      ) &
      all(is.na(loraz_equiv_total) | loraz_equiv_total == 0) &
      all(is.na(dexmedetomidine) | dexmedetomidine == 0),
      
      any_mme = any(!is.na(mme_total))
  ),
  by = hospitalization_id
][
  , opioid_only_analgesia := fifelse(all_sedatives_na_or_zero & any_mme, 1, 0)
]
opioid_only_summary[, .(hospitalization_id, opioid_only_analgesia)]

#### SENSITIVITY SECTION ###
total_timeline_hourly_primary <- as.data.frame(total_timeline_hourly_primary)

# Define columns to plot
columns_to_plot <- c("rass_value", "map", "spo2", "bilirubin_total", "creatinine", "platelet_count", 
                     "po2_arterial", "pco2_arterial", "ph_arterial", "fio2_combined", "gcs_total",
                     "dopamine","dobutamine", "norepinephrine", "epinephrine", "isoproterenol", "milrinone",
                     "phenylephrine", "rocuronium", "vasopressin", "vecuronium", "pao2_imputed")

# Convert data to long format
total_timeline_hourly_long <- total_timeline_hourly_primary %>%
  select_at(vars(one_of(columns_to_plot))) %>% 
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Plot histograms
histogram_plot <- ggplot(total_timeline_hourly_long, aes(x = Value)) +
  geom_histogram(fill = "steelblue", color = "black", bins = 30) +
  facet_wrap(~ Variable, scales = "free") +  # Adjust scales for continuous variables
  theme_minimal() +
  labs(title = "Histograms of Variables", x = "Value", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Save the Histogram
ggsave(
  filename = file.path(final_data, paste0("histograms_", institution, ".png")),
  plot = histogram_plot,
  width = 10,
  height = 6,
  dpi = 300
)

#Sensitivity Analysis Numbers
cols_to_update_imv_sensitivity <- c("hour_imv", "rass_value")

total_timeline_sensitivity <- merge(total_timeline_filt, time_dependent_timeline_last_sensitivity, by = c("hospitalization_id", "recorded_dttm", "hour_imv", "rass_value", "map", "spo2", "platelet_count", "po2_arterial", "ph_arterial", "gcs_total", "bilirubin_total", "creatinine", "pco2_arterial", "fio2_combined", "dopamine", "dobutamine", "norepinephrine", "epinephrine"), all = TRUE)

total_timeline_sensitivity[, pao2_imputed := {
  s <- spo2 / 100
  a <- 11700 / ((1 / s) - 1)
  b <- sqrt(50^3 + a^2)
  (b + a)^(1/3) - (b - a)^(1/3)
}]

total_timeline_sensitivity[, (cols_to_update_imv_sensitivity) := lapply(.SD, function(x) fifelse(x == "", NA, x)), .SDcols = cols_to_update_imv_sensitivity]

# Paralytics
total_timeline_sensitivity[, paralytics := fifelse(rocuronium > 0 | vecuronium > 0.09 | cisatracurium > 0.49, TRUE, FALSE)]
total_timeline_sensitivity[is.na(paralytics), paralytics := FALSE]

#Fix Logically Inconsistent RASS Values During Paralytic Periods
total_timeline_sensitivity[
  rass_value > -3 & paralytics == "TRUE",
  rass_value := NA
]
###

total_timeline_locf_sensitivity <- total_timeline_sensitivity[, (cols_to_update_imv_sensitivity) := lapply(.SD, function(x) na.locf(x, na.rm = FALSE)), 
                 by = hospitalization_id, .SDcols = cols_to_update_imv_sensitivity]

# Filter the table to retain rows where recorded_dttm is within the time range
total_timeline_locf_sensitivity_filt <- total_timeline_locf_sensitivity[recorded_dttm >= start_time & recorded_dttm <= end_time]

# Compute the min and max for each hospitalization_id and hour_imv
total_timeline_hourly_sensitivity <- total_timeline_locf_sensitivity_filt[, 
    lapply(.SD, min), by = .(hospitalization_id, hour_imv), .SDcols = cols_min][
    total_timeline_locf_sensitivity_filt[, lapply(.SD, max), by = .(hospitalization_id, hour_imv), .SDcols = cols_max], 
    on = .(hospitalization_id, hour_imv)
]

setDT(total_timeline_hourly_sensitivity)

total_timeline_hourly_sensitivity <- total_timeline_hourly_sensitivity[, .(hospitalization_id, hour_imv, rass_value)]
```

# PART SEVENTEEN: Calculate SOFA, Flag Paralysis, Merge with time independent table, calculate percent deep sedation
```{r}
setDT(total_timeline_hourly_primary)
# Paralytics
total_timeline_hourly_primary[, paralytics := fifelse(rocuronium > 0 | vecuronium > 0.09 | cisatracurium > 0.49, TRUE, FALSE)]
total_timeline_hourly_primary[is.na(paralytics), paralytics := FALSE]

# Calculate P/F Ratio
#total_timeline_hourly[, fio2_set := fifelse(fio2_set > 1, fio2_set / 100, fio2_set)]
total_timeline_hourly_primary[, p_to_f := po2_arterial / fio2_combined]
total_timeline_hourly_primary[, p_to_f_imputed := pao2_imputed / fio2_combined]

# Calculate SOFA
total_timeline_hourly_primary <- total_timeline_hourly_primary %>%
  mutate(
    sofa_cv_97 = case_when(
      dopamine > 15 | epinephrine > 0.1 | norepinephrine > 0.1 ~ 4,
      dopamine > 5 | (epinephrine <= 0.1 & epinephrine > 0) | (norepinephrine <= 0.1 & norepinephrine > 0) ~ 3,
      (dopamine <= 5 & dopamine > 0) | dobutamine > 0 ~ 2,
      map < 70 ~ 1,
      TRUE ~ 0
    ),
    sofa_coag = case_when(
      platelet_count < 20 ~ 4,
      platelet_count < 50 ~ 3,
      platelet_count < 100 ~ 2,
      platelet_count < 150 ~ 1,
      TRUE ~ 0
    ),
    sofa_liver = case_when(
      bilirubin_total >= 12.0 ~ 4,
      bilirubin_total >= 6.0 ~ 3,
      bilirubin_total >= 2.0 ~ 2,
      bilirubin_total >= 1.2 ~ 1,
      TRUE ~ 0
    ),
    sofa_renal = case_when(
      creatinine >= 5.0 ~ 4,
      creatinine >= 3.5 & creatinine < 5.0 ~ 3,
      creatinine >= 2.0 & creatinine < 3.5 ~ 2,
      creatinine >= 1.2 & creatinine < 2.0 ~ 1,
      TRUE ~ 0
    ),
    sofa_resp_pf = case_when(
      p_to_f < 100 ~ 4,
      p_to_f < 200 ~ 3,
      p_to_f < 300 ~ 2,
      p_to_f < 400 ~ 1,
      TRUE ~ 0
    ),
    sofa_resp_pf_imputed = case_when(
      p_to_f_imputed < 100 ~ 4,
      p_to_f_imputed < 200 ~ 3,
      p_to_f_imputed < 300 ~ 2,
      p_to_f_imputed < 400 ~ 1,
      TRUE ~ 0
    ),
    sofa_cns = case_when(
      gcs_total < 6 ~ 4,
      gcs_total >= 6 & gcs_total <= 9 ~ 3,
      gcs_total >= 10 & gcs_total <= 12 ~ 2,
      gcs_total >= 13 & gcs_total <= 14 ~ 1,
      TRUE ~ 0
    )
  )

# Compute the total SOFA score
total_timeline_hourly_primary <- copy(total_timeline_hourly_primary)
total_timeline_hourly_primary[, total_sofa := 
  sofa_cv_97 + sofa_coag + sofa_liver + sofa_renal +
  fcoalesce(sofa_resp_pf, sofa_resp_pf_imputed) +
  sofa_cns
]

max_sofa_by_hosp <- total_timeline_hourly_primary %>%
  group_by(hospitalization_id) %>%
  summarise(max_total_sofa = max(total_sofa, na.rm = TRUE), .groups = "drop")

# Deep Sedation Flag
total_timeline_hourly_rass_filter <- total_timeline_hourly_primary[, 
  deep_sedation_hourly := fifelse(is.na(rass_value), NA, fifelse(rass_value <= -3, TRUE, FALSE))
]

# Merge with Time Independent Table
independent_dependent_merge <- merge(total_timeline_hourly_rass_filter, final_cohort_imv_runs_reduced, by = "hospitalization_id", all = TRUE)

# Determine Continuous Paralytics Started at Intubation
time_to_first_paralytic <- independent_dependent_merge[
  hour_imv > 4, 
  .(time_to_first_paralytic = if (any(paralytics == TRUE)) min(hour_imv[paralytics == TRUE]) else NA_real_),
  by = hospitalization_id
]

# Determine Maximum SOFA
baseline_sofa <- max_sofa_by_hosp

#Calculate Percent Deep Sedation for total 72 hours
timeline_72h_no_first_four <- independent_dependent_merge %>%
  filter(hour_imv < 0 | hour_imv > 4)

deep_sedation_count_total_72 <- timeline_72h_no_first_four[
  deep_sedation_hourly == TRUE, 
  .N, 
  by = hospitalization_id
]
setnames(deep_sedation_count_total_72, "N", "deep_sedation_count_total_72")
deep_sedation_count_total_72[, deep_sedation_count_total_72 := deep_sedation_count_total_72]


row_count_minus_one_72 <- timeline_72h_no_first_four[
  , .(hours_mechanically_ventilated_72 = .N), 
  by = hospitalization_id
]

sedation_and_imv_72 <- merge(row_count_minus_one_72, deep_sedation_count_total_72, by = "hospitalization_id", all = TRUE)

# Replace NA values in deep_sedation_count_total_72 with 0
setnafill(sedation_and_imv_72, cols = "deep_sedation_count_total_72", fill = 0)

sedation_and_imv_72[, percent_deep_sedation_72 := (deep_sedation_count_total_72 / hours_mechanically_ventilated_72) * 100]


#Calculate Percent Deep Sedation for First 48 Hours
independent_dependent_merge_48 <- independent_dependent_merge[hour_imv >= 1 & hour_imv <= 48]

timeline_48h_no_first_four <- independent_dependent_merge_48 %>%
  filter(hour_imv < 0 | hour_imv > 4)

deep_sedation_count_total_48 <- timeline_48h_no_first_four[
  deep_sedation_hourly == TRUE, 
  .N, 
  by = hospitalization_id
]
setnames(deep_sedation_count_total_48, "N", "deep_sedation_count_total_48")
deep_sedation_count_total_48[, deep_sedation_count_total_48 := deep_sedation_count_total_48]


row_count_48 <- timeline_48h_no_first_four[
  , .(hours_mechanically_ventilated_48 = .N), 
  by = hospitalization_id
]

sedation_and_imv_48 <- merge(row_count_48, deep_sedation_count_total_48, by = "hospitalization_id", all = TRUE)

# Replace NA values in deep_sedation_count_total_48 with 0
setnafill(sedation_and_imv_48, cols = "deep_sedation_count_total_48", fill = 0)

sedation_and_imv_48[, percent_deep_sedation_48 := (deep_sedation_count_total_48 / hours_mechanically_ventilated_48) * 100]

##########
#Negative 1 Imputation Sensitivity Analysis
# Deep Sedation Flag
total_timeline_hourly_rass_sensitivity <- total_timeline_hourly_sensitivity[, 
  deep_sedation_hourly := fifelse(is.na(rass_value), NA, fifelse(rass_value <= -3, TRUE, FALSE))
]

#Calculate Percent Deep Sedation
total_timeline_hourly_rass_sensitivity_48 <- total_timeline_hourly_rass_sensitivity[hour_imv >= 0 & hour_imv <= 48]

timeline_48h_no_first_four_sensitivity <- total_timeline_hourly_rass_sensitivity_48 %>%
  filter(hour_imv < 0 | hour_imv > 4)

deep_sedation_count_total_48_sensitivity <- timeline_48h_no_first_four_sensitivity[
  deep_sedation_hourly == TRUE, 
  .N, 
  by = hospitalization_id
]
setnames(deep_sedation_count_total_48_sensitivity, "N", "deep_sedation_count_total_48_sensitivity")
deep_sedation_count_total_48_sensitivity[, deep_sedation_count_total_48_sensitivity := deep_sedation_count_total_48_sensitivity - 1]


row_count_minus_one_48_sensitivity <- timeline_48h_no_first_four_sensitivity[
  , .(hours_mechanically_ventilated_48_sensitivity = .N), 
  by = hospitalization_id
]

sedation_and_imv_48_sensitivity <- merge(row_count_minus_one_48_sensitivity, deep_sedation_count_total_48_sensitivity, by = "hospitalization_id", all = TRUE)

# Replace NA values in deep_sedation_count_total_48 with 0
setnafill(sedation_and_imv_48_sensitivity, cols = "deep_sedation_count_total_48_sensitivity", fill = 0)

sedation_and_imv_48_sensitivity[, percent_deep_sedation_48_sensitivity := (deep_sedation_count_total_48_sensitivity / hours_mechanically_ventilated_48_sensitivity) * 100]

sedation_and_imv_48_sensitivity[, .(hospitalization_id, percent_deep_sedation_48_sensitivity)]

##########

#Time to first rass

time_to_first_rass_analysis <- time_to_first_rass_final[hospitalization_id %in% final_cohort_imv_runs_reduced$hospitalization_id]

# Complete Time_independent_table for primary analysis
# List of tables to merge
primary_analysis_tables <- list(final_cohort_imv_runs_reduced, 
               time_to_first_paralytic,
               baseline_sofa, 
               sedation_and_imv_72,
               sedation_and_imv_48,
               sedation_and_imv_48_sensitivity,
               time_to_first_rass_analysis,
               median_mme_by_hosp,
               median_benzo_by_hosp,
               median_propofol_by_hosp,
               opioid_only_summary)

# Iterative Merge on Hospitalization_id to form primary analysis analytic
primary_analysis_raw <- reduce(primary_analysis_tables, function(x, y) merge(x, y, by = "hospitalization_id", all = TRUE))
# Replace NA values in rass_below_minus_3_count with 0
setnafill(primary_analysis_raw, cols = "rass_below_minus_3_count", fill = 0)

primary_analysis_raw_2 <- merge(primary_analysis_raw, paralytic_pushes_first_48h_final, by = "hospitalization_id", all = TRUE)
setnafill(primary_analysis_raw_2, cols = "push_paralytics", fill = 0)

primary_analysis_analytic <- primary_analysis_raw_2[, .(hospitalization_id, begin_imv, end_imv, admission_dttm, discharge_dttm, hospital_id, icu_type, hospital_icu_id, location_of_intubation, first_location_in_hospital, first_trach_dttm, imv_duration_hours_final, imv_duration_days, age_at_admission, age_category, death_dttm, admission_type_category, discharge_category, sex_category, race_category, ethnicity_category, ethnicity_category_final, SVM_SCORE, language_category, top_five_non_eng_non_spanish, non_eng, elixhauser_score, weight_kg, bmi, trached, reintubation, thirty_day_mortality, vfd_clean, hfd_6m, laps2, deep_sedation_count_total_48, deep_sedation_weighted, deep_sedation_unweighted, rass_weighted_avg, rass_below_minus_3_count, death_during_admission, max_total_sofa, time_to_first_paralytic, time_to_first_rass, hours_mechanically_ventilated_72, percent_deep_sedation_48, percent_deep_sedation_72, percent_deep_sedation_48_sensitivity, covid_status, rass_frequency_hrs, icufd_28, mean_gcs, median_gcs, median_rass_binary, median_rass_cont, sedation_index_score, time_to_icu_admission, hfd_28d, push_paralytics, loraz_median, mme_median, propofol_median, opioid_only_analgesia)]

write_parquet(primary_analysis_analytic, paste0(intermediate_dir, "analytic_file", institution, ".parquet"))
 
```

# PART EIGHTEEN: Data Cleaning Summary
```{r}
data_cleaning_summary <- data.table(
  Stage = c(
    "Total Hospitalizations", 
    "Documented hospitalizations with IMV",
    "Removed (Repeat episodes of IMV)",
    "Removed (IMV < 24 hours)",
    "Removed (Patients with Trach)",
    "Removed (Intubated at OSH)",
    "Removed (>4 hours between intubation and ICU transfer)",
    "Removed (Missing hospital or ICU ID info)",
    "Removed (Missing BMI Data)",
    "Removed (Missing demographic data or unknown/non-verbal language)",
    "Removed (Missing Elixhauser Data)",
    "Removed (No RASS within 8 hours of IMV)",
    "Removed (No RASS every 12 hours of IMV)",
    "Final Cohort"
  ),
  Table_Name = c(
    "hospitalization", 
    "imv_hospitalizations",
    "reintubations",
    "first_imv_runs - part_two_imv_runs",
    "part_two_imv_runs - part_three_imv_runs",
    "part_three_imv_runs - part_four_imv_runs",
    "part_four_imv_runs - location_of_intubation_final",
    "location_of_intubation_final - part_five_imv_runs",
    "part_five_imv_runs - part_seven_imv_runs",
    "part_seven_imv_runs - part_eight_imv_runs",
    "part_eight_imv_runs - part_nine_imv_runs",
    "part_nine_imv_runs - rass_cohort_timestamps",
    "rass_cohort_timestamps - part_ten_imv_runs",
    "part_eleven_imv_runs"
  ),
  Row_Count = c(
    nrow(hospitalization), 
    nrow(imv_hospitalizations),
    nrow(reintubations),
    nrow(first_imv_runs) - nrow(part_two_imv_runs),
    nrow(part_two_imv_runs) - nrow(part_three_imv_runs),
    nrow(part_three_imv_runs) - nrow(part_four_imv_runs),
    nrow(part_four_imv_runs) - nrow(location_of_intubation_final),
    nrow(location_of_intubation_final) - nrow(part_five_imv_runs),
    nrow(part_five_imv_runs) - nrow(part_seven_imv_runs),
    nrow(part_seven_imv_runs) - nrow(part_eight_imv_runs),
    nrow(part_eight_imv_runs) - nrow(part_nine_imv_runs),
    nrow(part_nine_imv_runs) - nrow(rass_cohort_timestamps),
    nrow(rass_cohort_timestamps) - nrow(part_ten_imv_runs),
    nrow(part_eleven_imv_runs)
  )
)

fwrite(data_cleaning_summary, file = paste0(final_data, institution, "_cohort_data_cleaning_summary.csv"))
```
